



\chapter{BB-SNAPSHOT.CGI}

\section{BB-SNAPSHOT.CGI}
 Section: User Commands (1) 
Updated: Version Exp: 11 Jan 2008 
Index Return to Main Contents 
 \section{NAME}
 bb-snapshot.cgi - CGI program to rebuild the Hobbit webpages for a specific point in time.  \section{SYNOPSIS}
\textbf{bb-snapshot.cgi}


  
\section{DESCRIPTION}
\textbf{bb-snapshot.cgi}
 is invoked as a CGI script via the bb-snapshot.sh CGI wrapper. It rebuilds the Hobbit web pages to the look they had at a particular point in time, based upon the historical information logged about events. 

  bb-snapshot.cgi is passed a QUERY\_STRING environment variable with the following parameters: 


  
   mon (Start month of the snapshot)  
   day (Start day-of-month of the snapshot)  
   yr  (Start year of the snapshot)  
   hour (Start hour of the snapshot)  
   min  (Start minute of the snapshot)  
   sec  (Start second of the snapshot) 


  The ``month'' parameters must be specified as the three-letter english month name abbreviation: Jan, Feb, Mar ... 


  ``day'' must be in the range 1..31; ``yr'' must be specified including century (e.g. ``2003''). ``hour'' must be specified using a 24-hour clock. 


  All of the processing involved in generating the report is done by invoking \emph{bbgen(1)}
 with the proper ``--snapshot'' option. 


  
\section{OPTIONS}
\begin{description}
\item[--env=FILENAME] Load environment from FILENAME before executing the CGI. 

 

\item[bbgen-options] All options except ``--env'' are passed on to the \emph{bbgen(1)}
 program building the snapshot files. 

 


\end{description}
 \section{ENVIRONMENT VARIABLES}
\begin{description}
\item[BBGENSNAPOPTS] bbgen options passed by default to the bb-snapshot.cgi script. This happens in the bb-snapshot.sh CGI wrapper script. 
\item[BBHOME] Home directory of the Hobbit server files 
\item[BBSNAP] Directory where generated snapshots are stored. This directory must be writable by the userid executing the CGI script, typically ``www'', ``apache'' or ``nobody''. Default: \$BBHOME/www/snap/ 
\item[BBSNAPURL] The URL prefix to use when accessing the reports via a browser. Default: \$BBWEB/snap 

 


 


\end{description}
 \section{SEE ALSO}
bbgen(1), bb-hosts(5), hobbitserver.cfg(5) 

 


  
 
\section{Index}
\begin{description}
\item[NAME]
\item[SYNOPSIS]
\item[DESCRIPTION]
\item[OPTIONS]
\item[ENVIRONMENT VARIABLES]
\item[SEE ALSO]

\end{description}
 
 This document was created by man2html, using the manual pages. 
 Time: 16:21:47 GMT, January 11, 2008 

