



\chapter{BB-SERVICES}

\section{BB-SERVICES}
 Section: File Formats (5) 
Updated: Version Exp: 11 Jan 2008 
Index Return to Main Contents 
 \section{NAME}
 bb-services - Configuration of TCP network services 

  
\section{SYNOPSIS}
\textbf{\$BBHOME/etc/bb-services}


  
\section{DESCRIPTION}
\textbf{bb-services}
 contains definitions of how \emph{bbtest-net(1)}
 should test a TCP-based network service (i.e. all common network services except HTTP and DNS). For each service, a simple dialogue can be defined to check that the service is functioning normally, and optional flags determine if the service has e.g. a banner or requires SSL- or telnet-style handshaking to be tested. 

  
\section{FILE FORMAT}
 bb-services is a text file. A simple service definition for the SMTP service would be this:  


  
   [smtp]  
 
      send ''mail$\backslash$r$\backslash$nquit$\backslash$r$\backslash$n''  
 
      expect ''220''  
 
      options banner  



  This defines a service called ``smtp''. When the connection is first established, bbtest-net will send the string ``mail$\backslash$r$\backslash$nquit$\backslash$r$\backslash$n'' to the service. It will then expect a response beginning with ``220''. Any data returned by the service (a so-called ``banner'') will be recorded and included in the status message. 


  The full set of commands available for the bb-services file are: 


 \begin{description}
\item[[NAME]] Define the name of the TCP service, which will also be the column-name in the resulting display on the test status. If multiple tests share a common definition (e.g. ssh, ssh1 and ssh2 are tested identically), you may list these in a single ``[ssh|ssh1|ssh2]'' definition, separating each service-name with a pipe-sign. 

 

\item[send STRING]
\item[expect STRING] Defines the strings to send to the service after a connection is established, and the response that is expected. Either of these may be omitted, in which case \emph{bbtest-net(1)}
 will simply not send any data, or match a response against anything. 

  The send- and expect-strings use standard escaping for non-printable characters. ``$\backslash$r'' represents a carriage-return (ASCII 13), ``$\backslash$n'' represents a line-feed (ASCII 10), ``$\backslash$t'' represents a TAB (ASCII 8). Binary data is input as ``$\backslash$xNN'' with NN being the hexadecimal value of the byte. 


 

\item[port NUMBER] Define the default TCP port-number for this service. If no portnumber is defined, \emph{bbtest-net(1)}
 will attempt to lookup the portnumber in the standard /etc/services file. 

 

\item[options option1[,option2][,option3]] Defines test options. The possible options are  
 
   banner - include received data in the status message  
 
   ssl - service uses SSL so perform an SSL handshake  
 
   telnet - service is telnet, so exchange telnet options 

 


 


\end{description}
 
\section{FILES}
\textbf{\$BBHOME/etc/bb-services}


  
\section{SEE ALSO}
bbtest-net(1) 

 


  
 
\section{Index}
\begin{description}
\item[NAME]
\item[SYNOPSIS]
\item[DESCRIPTION]
\item[FILE FORMAT]
\item[FILES]
\item[SEE ALSO]

\end{description}
 
 This document was created by man2html, using the manual pages. 
 Time: 16:21:46 GMT, January 11, 2008 

