\chapter{BBCOMBOTEST}

\section{NAME}
 bbcombotest - Hobbit combination test tool  \

section{SYNOPSIS}
\textbf{bbcombotest --help}
 
\textbf{bbcombotest --version}
 
\textbf{bbcombotest [--debug] [--quiet]}


  
\section{DESCRIPTION}
\textbf{bbcombotest} is a Hobbit extension script that runs on the
BBDISPLAY server. It combines the results of one or more of the normal
Hobbit test results into a combined test result, using standard
arithmetic og logical operators. 


  The resulting tests are sent to the Hobbit display server as any
  normal test - so all of the standard Hobbit functions (history,
  statistics etc.) are available for the combined tests. 



  The tool was born from the need to monitor systems with built-in
  redundancy and automatic failover - e.g. load-balanced web
  servers. But other uses are possible. 



  
\section{OPTIONS}
\begin{description}
\item[--quiet] Normally, the test status sent by bbcombotest includes
  information about the underlying test results used to determine the
  current value of the combined test. ``--quiet'' eliminates this
  information from the test status page. 


 

\item[--debug] Provide debugging output for use in troubleshooting problems with bbcombotest. 

 

\item[--no-update] Dont send any status messages - instead, the result of the combotests is simply dumped to stdout. Useful for debugging. 


\end{description}
 \section{FILES}
\begin{description}
\item[\$BBHOME/etc/bbcombotest.cfg] Configuration file for bbcombotest, where the combined tests are defined 
\item[\$BBHOME/etc/hobbitlaunch.cfg] Configuration file controlling when bbcombotest is run. 


\end{description}
 \section{SEE ALSO}
bbcombotest.cfg(5), bb-hosts(5), hobbitserver.cfg(5), hobbitlaunch.cfg(5) 


  

