



\chapter{BBHOSTGREP}

\section{BBHOSTGREP}
 Section: User Commands (1) 
Updated: Version Exp: 11 Jan 2008 
Index Return to Main Contents 
 \section{NAME}
 bbhostgrep - pick out lines in bb-hosts  \section{SYNOPSIS}
\textbf{bbhostgrep --help}
 
\textbf{bbhostgrep --version}
 
\textbf{bbhostgrep [--noextras] [--test-untagged] [--bbdisp] [--bbnet] TAG [TAG...]}


  
\section{DESCRIPTION}
\emph{bbhostgrep(1)}
 is for use by extension scripts that need to pick out the entries in a bb-hosts file that are relevant to the script. 

  The utility accepts test names as parameters, and will then parse the bb-hosts file and print out the host entries that have at least one of the wanted tests specified. Tags may be given with a trailing asterisk '*', e.g. ``bbhostgrep http*'' is needed to find all http and https tags. 


  The bbhostgrep utility supports the use of ``include'' directives inside the bb-hosts file, and will find matching tags in all included files. 


  If the DOWNTIME or SLA tags are used in the \emph{bb-hosts(5)}
 file, these are interpreted relative to the current time. bbhostgrep then outputs a ``INSIDESLA'' or ``OUTSIDESLA'' tag for easier use by scripts that want to check if the current time is inside or outside the expected uptime window. 


  
\section{OPTIONS}
\begin{description}
\item[--noextras] Remove the ``testip'', ``dialup'', ``INSIDESLA'' and ``OUTSIDESLA'' tags from the output. 

 

\item[--test-untagged] When using the BBLOCATION environment variable to test only hosts on a particular network segment, bbtest-net will ignore hosts that do not have any ``NET:x'' tag. So only hosts that have a NET:\$BBLOCATION tag will be tested.  
 With this option, hosts with no NET: tag are included in the test, so that all hosts that either have a matching NET: tag, or no NET: tag at all are tested. 

 

\item[--no-down[=TESTNAME]] bbhostgrep will query the Hobbit server for the current status of the ``conn'' test, and if TESTNAME is specified also for the current state of the specified test. If the status of the ``conn'' test for a host is non-green, or the status of the TESTNAME test is disabled, then this host is ignored and will not be included in the output. This can be used to ignore hosts that are down, or hosts where the custom test is disabled. 

 

\item[--bbdisp] Search the bb-hosts file following include statements as a BBDISPLAY server would. 

 

\item[--bbnet] Search the bb-hosts file following include statements as a BBNET server would. 

 


\end{description}
 \section{EXAMPLE}
 If your bb-hosts file looks like this 

  
   192.168.1.1   www.test.com  \# ftp telnet !oracle  
   192.168.1.2   db1.test.com  \# oracle  
   192.168.1.3   mail.test.com \# smtp 


  and you have a custom Hobbit extension script that performs the ``oracle'' test, then running ``bbhostgrep oracle'' would yield 


  
   192.168.1.1   www.test.com  \# !oracle  
   192.168.1.2   db1.test.com  \# oracle 


  so the script can quickly find the hosts that are of interest. 


  Note that the reverse-test modifier - ``!oracle'' - is included in the output; this also applies to the other test modifiers defined by Hobbit (the dialup and always-true modifiers). 


  If your extension scripts use more than one tag, just list all of the interesting tags on the command line. 


  bbhostgrep also supports the ``NET:location'' tag used by bbtest-net, so if your script performs network checks then it will see only the hosts that are relevant for the test location that the script currently executes on. 


  
\section{USE IN EXTENSION SCRIPTS}
 To integrate bbhostgrep into an existing script, look for the line in the script that grep's in the \$BBHOSTS file. Typically it will look somewhat like this: 

  
   \$GREP -i ''\^{}[0-9].*\#.*TESTNAME'' \$BBHOSTS | ... code to handle test 


  Instead of the grep, we will use bbhostgrep. It then becomes 


  
   \$BBHOME/bin/bbhostgrep TESTNAME | ... code to handle test 


  which is simpler, less error-prone and more efficient. 


  
\section{ENVIRONMENT VARIABLES}
\begin{description}
\item[BBLOCATION] If set, bbhostgrep outputs only lines from bb-hosts that have a matching NET:\$BBLOCATION setting. 

 

\item[BBHOSTS] Filename for the Hobbit \emph{bb-hosts(5)}
 file. 

 


\end{description}
 \section{FILES}
\begin{description}
\item[\$BBHOSTS] The Hobbit bb-hosts file 

 


\end{description}
 \section{SEE ALSO}
bb-hosts(5), hobbitserver.cfg(5) 

  
 
\section{Index}
\begin{description}
\item[NAME]
\item[SYNOPSIS]
\item[DESCRIPTION]
\item[OPTIONS]
\item[EXAMPLE]
\item[USE IN EXTENSION SCRIPTS]
\item[ENVIRONMENT VARIABLES]
\item[FILES]
\item[SEE ALSO]

\end{description}
 
 This document was created by man2html, using the manual pages. 
 Time: 16:21:46 GMT, January 11, 2008 

