



\chapter{BBMESSAGE.CGI}

\section{BBMESSAGE.CGI}
 Section: Maintenance Commands (8) 
Updated: Version Exp: 11 Jan 2008 
Index Return to Main Contents 


  
\section{NAME}
 bbmessage.cgi - CGI utility used for proxying Hobbit data over HTTP  \section{SYNOPSIS}
\textbf{bbmessage.cgi}


  
\section{DESCRIPTION}
\emph{bbmessage.cgi(8)}
 is the server-side utility receiving Hobbit messages sent by the \emph{bb(1)}
 utility over an HTTP transport. The \textbf{bb}
 utility normally sends data over a dedicated TCP protocol, but it may use HTTP to go through proxies or through restrictive firewalls. In that case, the webserver must have this CGI utility installed, which takes care of receiving the message via HTTP, and forwards it to a local Hobbit server through the normal Hobbit transport. 

  The CGI expects to be invoked from an HTTP ``POST'' request, with the POST-data being the status-message. \textbf{bbmessage.cgi}
 simply collects all of the POST data, and send it off as a message to the Hobbit daemon running on IP 127.0.0.1. This destination IP currently cannot be changed. 


  The CGI will return any output provided by the Hobbit daemon back to the requestor as the response to the HTTP POST, so this allows for all normal Hobbit commands to work. 


  
\section{SECURITY}
\textbf{bbmesage.cgi}
 will only send data to a Hobbit server through the loopback interface, i.e. IP-address 127.0.0.1. 

  Access to the CGI should be restricted through webserver access controls, since the CGI provides no authentication at all to validate incoming messages. 


  If possible, consider using the \emph{bbproxy(8)}
 utility instead for native proxying of Hobbit data between networks. 


  
\section{SEE ALSO}
bb(1), bbproxy(8), hobbit(7) 

 


  
 
\section{Index}
\begin{description}
\item[NAME]
\item[SYNOPSIS]
\item[DESCRIPTION]
\item[SECURITY]
\item[SEE ALSO]

\end{description}
 
 This document was created by man2html, using the manual pages. 
 Time: 16:21:46 GMT, January 11, 2008 

