



\chapter{HOBBIT-STATUSREPORT.CGI}

\section{HOBBIT-STATUSREPORT.CGI}
 Section: User Commands (1) 
Updated: Version Exp: 11 Jan 2008 
Index Return to Main Contents 
 \section{NAME}
 hobbit-statusreport.cgi - CGI program to report a status for a group of servers  \section{SYNOPSIS}
\textbf{hobbit-statusreport.cgi --column=COLUMNNAME [options]}


  
\section{DESCRIPTION}
\textbf{hobbit-statusreport.cgi}
 is a CGI tool to generate a simple HTML report showing the current status of a single column for a group of Hobbit hosts. 

  E.g. You can use this report to get an overview of all of the SSL certificates that are about to expire. 


  The generated webpage is a simple HTML table, suitable for copying into other documents or e-mail. 


  hobbit-statusreport.cgi runs as a CGI program, invoked by your webserver. It is normally run via a wrapper shell-script in the CGI directory for Hobbit. 


  
\section{EXAMPLES}
 The Hobbit installation includes two web report scripts using this CGI tool: The \textbf{hobbit-certreport.sh}
 script generates a list of SSL server certificates that are yellow or red (i.e. they will expire soon); and the \textbf{hobbit-nongreen.sh}
 script generates a report of all statuses that are currently non-green. These can be accessed from a web browser through a URL referencing the script in the Hobbit CGI directory (e.g. ``/hobbit-cgi/hobbit-nongreen.sh''). 

  
\section{OPTIONS}
\begin{description}
\item[--column=COLUMNNAME] Report the status of the COLUMNNAME column. 

 

\item[--all] Report the status for all hosts known to Hobbit. By default, this tool reports only on the hosts found on the current page from where the CGI was invoked (by looking at the ``pagepath'' cookie). 

 

\item[--filter=CRITERIA] Only report on statuses that match the CRITERIA setting. See the \emph{bb(1)}
 man-page - in the ``hobbitdboard'' command description - for details about specifying filters. 

 

\item[--heading=HTML] Defines the webpage heading - i.e. the ``title'' tag in the generated HTML code. 

 

\item[--show-column] Include the column name in the display. 

 

\item[--show-colors] Show the status color on the generated webpage. The default is to not show the status color. 

 

\item[--no-colors] Do not include text showing the current color of each status in the report. This is the default. 

 

\item[--show-summary] Show only a summary of the important lines in the status message. By default, the entire status message appears in the generated HTML code. This option causes the first non-blank line of the status message to be shown, and also any lines beginning with ``\&COLOR'' which is used by many status messages to point out lines of interest (non-green lines only, though). 

 

\item[--show-message] Show the entire message on the webpage. This is the default. 

 

\item[--link] Include HTML links to the host ``info'' page, and the status page. 

 

\item[--embedded] Only generate the HTML table, not a full webpage. This can be used to embed the status report into an external webpage. 

 

\item[--env=FILENAME] Load the environment from FILENAME before executing the CGI. 

 

\item[--area=NAME] Load environment variables for a specific area. NB: if used, this option must appear before any --env=FILENAME option. 

 


\end{description}
 \section{SEE ALSO}
hobbit(7) 

 


  
 
\section{Index}
\begin{description}
\item[NAME]
\item[SYNOPSIS]
\item[DESCRIPTION]
\item[EXAMPLES]
\item[OPTIONS]
\item[SEE ALSO]

\end{description}
 
 This document was created by man2html, using the manual pages. 
 Time: 16:21:47 GMT, January 11, 2008 

