



\chapter{HOBBIT-CONFREPORT.CGI}

\section{HOBBIT-CONFREPORT.CGI}
 Section: User Commands (1) 
Updated: Version Exp: 11 Jan 2008 
Index Return to Main Contents 
 \section{NAME}
 hobbit-confreport.cgi - Hobbit Configuration report  \section{SYNOPSIS}
\textbf{hobbit-confreport.cgi}


  
\section{DESCRIPTION}
\textbf{hobbit-confreport.cgi}
 is invoked as a CGI script via the hobbit-confreport.sh CGI wrapper. 

 \textbf{hobbit-confreport.cgi}
 provides a plain HTML (Web) report of the Hobbit configuration for a group of hosts; which hosts are included is determined by the hosts available on the webpage from where the CGI script is invoked. 


  The configuration report include the hostnames, a list of the statuses monitored for each host, and if applicable any configuration settings affecting these. Alerts that may be triggered by status changes are also included. 


  The report is plain HTML without any images included, and therefore suitable for inclusion into e-mails or other documents that may be accessed outside the Hobbit system. 


  
\section{OPTIONS}
\begin{description}
\item[--critical] Report only on the statuses that are configured to show up on the \textbf{Critical Systems}
 view. 

 

\item[--old-nk-config] Use the deprecated \textbf{NK}
 tag in bb-hosts to determine if tests appear on the Critical Systems view. 

 

\item[--env=FILENAME] Loads the environment defined in FILENAME before executing the CGI script. 

 

\item[--area=NAME] Load environment variables for a specific area. NB: if used, this option must appear before any --env=FILENAME option. 

 

\item[--debug] Enables debugging output. 

 


\end{description}
 \section{BUGS}
 Client-side configuration done in the \emph{hobbit-clients.cfg(5)}
 is not currently reflected in the report. 

  Critical Systems view configuration is not reflected in the report. 


  
\section{SEE ALSO}
bb-hosts(5), hobbit-alerts.cfg(5), hobbit-clients.cfg(5), hobbit(7) 

 


  
 
\section{Index}
\begin{description}
\item[NAME]
\item[SYNOPSIS]
\item[DESCRIPTION]
\item[OPTIONS]
\item[BUGS]
\item[SEE ALSO]

\end{description}
 
 This document was created by man2html, using the manual pages. 
 Time: 16:21:47 GMT, January 11, 2008 

