



\chapter{HOBBIT-NKEDIT.CGI}

\section{HOBBIT-NKEDIT.CGI}
 Section: User Commands (1) 
Updated: Version Exp: 11 Jan 2008 
Index Return to Main Contents 
 \section{NAME}
 hobbit-nkedit.cgi - Hobbit Critical Systems view CGI  \section{SYNOPSIS}
\textbf{hobbit-nkedit.cgi}


  
\section{DESCRIPTION}
\textbf{hobbit-nkedit.cgi}
 is invoked as a CGI script via the hobbit-nkedit.sh CGI wrapper. 

  hobbit-nkedit.cgi is a web-based editor for the \emph{hobbit-nkview.cfg(5)}
 file, which is used to configure the Hobbit ``Critical Systems'' view. 


  A detailed description of how to use the editor is provided in the Hobbit Web documentation, available from the ``Help'' -$>$ ``Critical Systems'' link on the Hobbit website. 


  
\section{SECURITY}
 Access to this CGI script should be restricted through access controls in your webserver. Editing the Critical Systems View configuration will impact the monitoring of your site. 

  
\section{OPTIONS}
\begin{description}
\item[--config=FILENAME] Name of the Critical Systems View configuration file. The default is hobbit-nkview.cfg in the \$BBHOME/etc/ directory. 

 

\item[--env=FILENAME] Loads the environment defined in FILENAME before executing the CGI script. 

 

\item[--area=NAME] Load environment variables for a specific area. NB: if used, this option must appear before any --env=FILENAME option. 

 

\item[--debug] Enables debugging output. 

 


\end{description}
 \section{SEE ALSO}
hobbit-nkview.cgi(1), hobbit(7) 

 


  
 
\section{Index}
\begin{description}
\item[NAME]
\item[SYNOPSIS]
\item[DESCRIPTION]
\item[SECURITY]
\item[OPTIONS]
\item[SEE ALSO]

\end{description}
 
 This document was created by man2html, using the manual pages. 
 Time: 16:21:47 GMT, January 11, 2008 

