



\chapter{HOBBITCGI.CFG}

\section{HOBBITCGI.CFG}
 Section: File Formats (5) 
Updated: Version Exp: 11 Jan 2008 
Index Return to Main Contents 
 \section{NAME}
 hobbitcgi.cfg - Command-line parameters for the Hobbit CGI tools 

  
\section{SYNOPSIS}
\textbf{\$BBHOME/etc/hobbitcgi.cfg}


  
\section{DESCRIPTION}
\emph{hobbitcgi.cfg(1)}
 controls the commandline options passed to all of the Hobbit CGI tools through their respective shell-script wrappers. Typically the options listed here are used for system-wide configuration of the CGI utilities, e.g. to define where they read configuration files. 

  The exact set of commandline options available are described in the man-page for each of the CGI utilities. 


  The file is ``sourced'' into the shell script wrapper, so assignments to the CGI-specific variables must follow standard shell-script syntax. 


  
\section{SETTINGS}
\begin{description}
\item[CGI\_ACKINFO\_OPTS] Options for the \emph{hobbit-ackinfo.cgi(1)}
 utility. 

 

\item[CGI\_ACK\_OPTS] Options for the \emph{bb-ack.cgi(1)}
 utility. 

 

\item[CGI\_CSVINFO\_OPTS] Options for the \emph{bb-csvinfo.cgi(1)}
 utility. 

 

\item[CGI\_DATEPAGE\_OPTS] Options for the \emph{bb-datepage.cgi(1)}
 utility. 

 

\item[CGI\_ENADIS\_OPTS] Options for the \emph{hobbit-enadis.cgi(8)}
 utility. 

 

\item[CGI\_EVENTLOG\_OPTS] Options for the \emph{bb-eventlog.cgi(1)}
 utility. 

 

\item[CGI\_FINDHOST\_OPTS] Options for the \emph{bb-findhost.cgi(1)}
 utility. 

 

\item[CGI\_HIST\_OPTS] Options for the \emph{bb-hist.cgi(1)}
 utility. 

 

\item[CGI\_HOBBITCOLUMN\_OPTS] Hobbit-specific options for column documentation. This uses the \emph{bb-csvinfo.cgi(1)}
 utility with the \textbf{server/etc/columndoc.cfg}
 configuration file. 

 

\item[CGI\_HOBBITCONFREPORT\_OPTS] Options for the \emph{hobbit-confreport.cgi(1)}
 utility. 

 

\item[CGI\_HOBBITGRAPH\_OPTS] Options for the \emph{hobbitgraph.cgi(1)}
 utility. 

 

\item[CGI\_HOSTGRAPHS\_OPTS] Options for the \emph{hobbit-hostgraphs.cgi(1)}
 utility. 

 

\item[CGI\_NKEDIT\_OPTS] Options for the \emph{hobbit-nkedit.cgi(1)}
 utility. 

 

\item[CGI\_NKVIEW\_OPTS] Options for the \emph{hobbit-nkview.cgi(1)}
 utility. 

 

\item[CGI\_REPLOG\_OPTS] Options for the \emph{bb-replog.cgi(1)}
 utility. 

 

\item[CGI\_REP\_OPTS] Options for the \emph{bb-rep.cgi(1)}
 utility. 

 

\item[CGI\_SNAPSHOT\_OPTS] Options for the \emph{bb-snapshot.cgi(1)}
 utility. 

 

\item[CGI\_SVCHIST\_OPTS] Options for the \emph{hobbitsvc.cgi(1)}
 utility when used to view historical logs. Note that the ``--historical'' option must be included in this setting. 

 

\item[CGI\_SVC\_OPTS] Options for the \emph{hobbitsvc.cgi(1)}
 utility. 

 


\end{description}
 \section{SEE ALSO}
hobbit(7), the individual CGI utility man-pages. 

 


  
 
\section{Index}
\begin{description}
\item[NAME]
\item[SYNOPSIS]
\item[DESCRIPTION]
\item[SETTINGS]
\item[SEE ALSO]

\end{description}
 
 This document was created by man2html, using the manual pages. 
 Time: 16:21:47 GMT, January 11, 2008 

