



\chapter{HOBBITCLIENT.CFG}

\section{HOBBITCLIENT.CFG}
 Section: File Formats (5) 
Updated: Version Exp: 11 Jan 2008 
Index Return to Main Contents 
 \section{NAME}
 hobbitclient.cfg - Hobbit client environment variables 

  
\section{DESCRIPTION}
 Hobbit programs use multiple environment variables beside the normal set of variables. For the Hobbit client, the environment definitions are stored in the ~hobbit/client/etc/hobbitclient.cfg file. Each line in this file is of the form \textbf{NAME=VALUE}
 and defines one environment variable NAME with the value VALUE. 

  
\section{SETTINGS}
\begin{description}
\item[BBDISP] The IP-address used to contact the Hobbit server. Default: Chosen when the Hobbit client was compiled. 

 

\item[BBDISPLAYS] List of IP-adresses of Hobbit servers. Data will be sent to all of the servers listed here. This setting is only used if BBDISP=0.0.0.0. 

 

\item[BBPORT] The portnumber for used to contact the Hobbit server. Default: 1984. 

 

\item[BBHOME] The Hobbit client top-level directory. Default: The \$HOBBITCLIENTHOME setting inherited from the ``runclient.sh'' script which starts the Hobbit client. 

 

\item[BBCLIENTLOGS] The directory for the Hobbit clients' own logfiles. Default: \$BBHOME/logs 

 

\item[BBTMP] Directory used for temporary files. Default: \$BBHOME/tmp/ 

 

\item[BB] Full path to the \emph{bb(1)}
 client program. Default: \$BBHOME/bin/bb. 

 

\item[Commands] Many extension scripts expect a series of environment variables to point at various system utilities. These are included in the file when the client is built. 

 


\end{description}
 \section{INHERITED SETTINGS}
 Some environment variables are inherited from the ``runclient.sh'' script which launches the Hobbit client: 

 \begin{description}
\item[MACHINEDOTS] The hostname of the local system. Default: Taken from ``uname -n''. 

 

\item[MACHINE] The hostname of the local system, with dots replaced by commas. For compatibility with Big Brother extension scripts. 

 

\item[BBOSTYPE] The operating system of the local system, in lowercase. Default: taken from ``uname -s''. 

 

\item[HOBBITCLIENTHOME] The top-level directory for the Hobbit client. Default: The location of the ``runclient.sh'' script. 

 


\end{description}
 
\section{SEE ALSO}
hobbit(7) 

 


  
 
\section{Index}
\begin{description}
\item[NAME]
\item[DESCRIPTION]
\item[SETTINGS]
\item[INHERITED SETTINGS]
\item[SEE ALSO]

\end{description}
 
 This document was created by man2html, using the manual pages. 
 Time: 16:21:46 GMT, January 11, 2008 

