



\chapter{HOBBITD\_CLIENT}

\section{HOBBITD\_CLIENT}
 Section: Maintenance Commands (8) 
Updated: Version Exp: 11 Jan 2008 
Index Return to Main Contents 
 \section{NAME}
 hobbitd\_client - hobbitd worker module for client data  \section{SYNOPSIS}
\textbf{hobbitd\_channel --channel=client hobbitd\_client [options]}


  
\section{DESCRIPTION}
 hobbitd\_client is a worker module for hobbitd, and as such it is normally run via the \emph{hobbitd\_channel(8)}
 program. It receives hobbitd client messages sent from systems that have the the Hobbit client installed, and use the client data to generate the Hobbit status messages for the cpu-, disk-, memory- and procs-columns. It also feeds Hobbit data messages with the netstat- and vmstat-data collected by the client. 

  When generating these status messages from the client data, hobbitd\_client will use the configuration rules defined in the \emph{hobbit-clients.cfg(5)}
 file to determine the color of each status message. 


  
\section{OPTIONS}
\begin{description}
\item[--clear-color=COLOR] Define the color used when sending ``msgs'', ``files'' or ``ports'' reports and there are no rules to check for these statuses. The default is to show a ``clear'' status, but some people prefer to have it ``green''. If you would rather prefer not to see these status columns at all, then you can use the ``--no-clear-msgs'', ``--no-clear-files'' and ``--no-clear-ports'' options instead. 

 

\item[--no-clear-msgs] If there are no logfile checks, the ``msgs'' column will show a ``clear'' status. If you would rather avoid having a ``msgs'' column, this option causes hobbitd\_client to not send in a clear ``msgs'' status. 

 

\item[--no-clear-files] If there are no file checks, the ``files'' column will show a ``clear'' status. If you would rather avoid having a ``files'' column, this option causes hobbitd\_client to not send in a clear ``files'' status. 

 

\item[--no-clear-ports] If there are no port checks, the ``ports'' column will show a ``clear'' status. If you would rather avoid having a ``ports'' column, this option causes hobbitd\_client to not send in a clear ``ports'' status. 

 

\item[--no-ps-listing] Normally the ``procs'' status message includes the full process-listing received from the client. If you prefer to just have the monitored processes shown, this option will turn off the full ps-listing. 

 

\item[--no-port-listing] Normally the ``ports'' status message includes the full netstat-listing received from the client. If you prefer to just have the monitored ports shown, this option will turn off the full netstat-listing. 

 

\item[--config=FILENAME] Sets the filename for the \emph{hobbit-clients.cfg}
 file. The default value is ``etc/hobbit-clients.cfg'' below the Hobbit server directory. 

 

\item[--dump-config] Dumps the configuration after parsing it. May be useful to track down problems with configuration file errors. 

 

\item[--test] Starts an interactive session where you can test the hobbit-clients.cfg configuration. 

 

\item[--debug] Enable debugging output. 

 


\end{description}
 \section{FILES}
\begin{description}
\item[~hobbit/server/etc/hobbit-clients.cfg]

 


\end{description}
 \section{SEE ALSO}
hobbit-clients.cfg(5), hobbitd(8), hobbitd\_channel(8), hobbit(7) 

 


  
 
\section{Index}
\begin{description}
\item[NAME]
\item[SYNOPSIS]
\item[DESCRIPTION]
\item[OPTIONS]
\item[FILES]
\item[SEE ALSO]

\end{description}
 
 This document was created by man2html, using the manual pages. 
 Time: 16:21:46 GMT, January 11, 2008 

