%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Server Configuration files}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{hobbitserver.cfg - Hobbit Environment Variables }

 
\subsection{DESCRIPTION}
 Hobbit programs use multiple environment variables beside the normal
 set of variables. The environment definitions are stored in the
 ~hobbit/server/etc/hobbitserver.cfg file. Each line in this file is
 of the form \textbf{NAME=VALUE} and defines one environment variable
 NAME with the value VALUE. 


 
\subsection{ENVIRONMENT AREAS}
 In some cases it may be useful to have different values for an
 environment variable, depending on where it is used. This is possible
 by defining variables with an associated ``area''. Such definitions
 have the form \textbf{AREA/NAME=VALUE}. 


  E.g. to define a special setup of the BBDISPLAY variable when it is
  used by an application in the ``management'' area, you would do
  this: 

\begin{verbatim}

BBDISP="127.0.0.1"            # Default definition
management/BBDISP="10.1.0.5"  # Definition in the "management" area

\end{verbatim}



Areas are invoked by using the ``--area'' option for all tools, or via
the ENVAREA setting in the \emph{hobbitlaunch.cfg(5)} file. 


\subsection{GENERAL SETTINGS}


\begin{description}

\item[\index{BBSERVERHOSTNAME}] The fully-qualified hostname of the server that is running Hobbit. 

\item[\index{BBSERVERWWWNAME}] The hostname used to access this servers'
  web-pages, used to construct URL's in the Hobbit webpages. Default
  is the BBSERVERHOSTNAME. 


\item[\index{BBSERVERIP}] The public IP-address of the server that is running Hobbit. 
 

\item[\index{BBSERVEROS}] A name identifying the operating system of the
  Hobbit server. The known names are currently ``linux'', ``freebsd'',
  ``solaris'', ``hpux'', ``aix'' and ``osf''. 

\item[\index{FQDN}] If set to TRUE, Hobbit will use fully-qualified hostnames
  throughout. If set to FALSE, hostnames are stripped of their
  domain-part before being processed. It is \textbf{highly
  recommended} that you keep this set to TRUE. Default: TRUE. 

\item[\index{BBLOGSTATUS}] Controls how the HTML page for a status log is
  generated. If set to DYNAMIC, the HTML logs are generated on-demand
  by the \emph{bb-hostsvc.cgi(1)} script. If set to STATIC, you must
  activate the \emph{hobbitd\_filestore(8)} module (through an entry
  in the \emph{hobbitlaunch.cfg(5)} file) to create and store the HTML
  logs whenever a status update is received. Setting
  ``BBLOGSTATUS=STATIC'' is \textbf{discouraged} since the I/O load on
  the Hobbit server will increase significantly. 


\item[\index{PINGCOLUMN}] Defines the name of the column for ``ping'' test
  status. The data from the ``ping'' test is used internally by
  Hobbit, so it must be defined here so all of the Hobbit tools know
  which column to watch for this data. The default setting is
  PINGCOLUMN=conn. 

\item[\index{INFOCOLUMN}] Defines the name of the column for the ``info'' pages. 

\item[\index{TRENDSCOLUMN}] Defines the name of the column for the RRD graph pages. 

\item[\index{RRDHEIGHT}] The default height (in pixels) of the RRD graph images. Default: 120 pixels. 

\item[\index{RRDWIDTH}] The default width (in pixels) of the RRD graph images. Default: 576 pixels. 

\item[\index{TRENDSECONDS}] The graphs on the ``trends'' page show data for
  the past TRENDSECONDS seconds. Default: 172800 seconds, i.e. 48
  hours. 
 

\item[\index{HTMLCONTENTTYPE}] The Content-type reported by the CGI scripts
  that generate web pages. By default, this it ``text/html''. If you
  have on-line help texts in character sets other than the ISO-8859-1
  (western european) character set, it may be necessary to modify this
  to include a character set. E.g.  

 
HTMLCONTENTTYPE=''text/html;charset=euc-jp''   for a Japanese
character sets. Note: Some webservers will automatically add this, if
configured to do so. 



\item[\index{HOLIDAYS}] Defines the default set of holidays used if there is
  no ``holidays'' tag for a host in the bb-hosts file. Holiday sets
  are defined in the \emph{hobbit-holidays.cfg(5)} file. If not
  defined, only the default holidays (those defined outside a named
  holiday set) will be considered as holidays. 



\item[\index{WEEKSTART}] Defines which day is the first day of the week. Set
  to ``0'' for Sunday, ``1'' for Monday. Default: 1 (Monday). 
 

\end{description}

\subsection{DIRECTORIES}

\begin{description}

\item[\index{BBSERVERROOT}] The top-level directory for the Hobbit
  installation. The default is the home-directory for the user running
  Hobbit. 


 

\item[\index{BBSERVERLOGS}] The directory for the Hobbit's own logfiles (NOT
  the status-logs from the monitored hosts). 


 

\item[\index{BBHOME}] The Hobbit server directory, where programs and
  configurations are kept. Default: \$BBSERVERROOT/server/ . 


\item[\index{BBTMP}] Directory used for temporary files. Default: \$BBHOME/tmp/ 

 
\item[\index{BBWWW}] Directory for Hobbit webfiles. The \$BBWEB URL must map
  to this directory. Default: \$BBHOME/www/ 


\item[\index{BBNOTES}] Directory for Hobbit notes-files. The \$BBNOTESSKIN URL
  must map to this directory. Default: \$BBHOME/www/notes/ 



\item[\index{BBREP}] Directory for Hobbit availability reports. The \$BBREPURL
  URL must map to this directory. Note also that your webserver must
  have write-access to this directory, if you want to use the
  \emph{bb-rep.cgi(1)} CGI script to generate reports
  on-demand. Default: \$BBHOME/www/rep/ 


 

\item[\index{BBSNAP}] Directory for Hobbit snapshots. The \$BBSNAPURL URL must
  map to this directory. Note also that your webserver must have
  write-access to this directory, if you want to use the
  \emph{bb-snapshot.cgi(1)} CGI script to generate snapshots
  on-demand. Default: \$BBHOME/www/snap/ 


 

\item[\index{BBVAR}] Directory for all data stored about the monitored items. Default: \$BBSERVERROOT/data/ 

 

\item[\index{BBLOGS}] Directory for storing the raw status-logs. Not used
  unless ``hobbitd\_filestore --status'' is running, which is
  \textbf{discouraged} since it increases the load on the Hobbit
  server significantly. Default: \$BBVAR/logs/ 


 

\item[\index{BBHTML}] Directory for storing HTML status-logs. Not used unless
  ``hobbitd\_filestore --status --html'' is running, which is
  \textbf{discouraged} since it increases the load on the Hobbit
  server significantly. Default: \$BBHOME/www/html/ 


 

\item[\index{BBHIST}] Directory for storing the history of monitored items. Default: \$BBVAR/hist/ 

 

\item[\index{BBHISTLOGS}] Directory for storing the detailed status-log of
  historical events. Default: \$BBVAR/histlogs/ 


 

\item[\index{BBACKS}] Directory for storing information about alerts that have been acknowledged. Default: \$BBVAR/acks/ 

 

\item[\index{BBDISABLED}] Directory for storing information about tests that have been disabled. Default: \$BBVAR/disabled/ 

 

\item[\index{BBDATA}] Directory for storing incoming ``data'' messages. Default: \$BBVAR/data/ 

 

\item[\index{BBRRDS}] Top-level directory for storing RRD files (the databases
  with trend-information used to generate graphs). Default:
  \$BBVAR/rrd/ 


 

\item[\index{\index{CLIENTLOGS}}] Directory for storing the data sent by a Hobbit
  client around the time a status changes to a warning (yellow) or
  critical (red) state. Used by the \emph{hobbitd\_hostdata(8)}

 module. Default: \$BBVAR/hostdata/ 

 


 


\end{description}

\subsection{SYSTEM FILES}


 \begin{description}
\item[\index{BBHOSTS}] Full path to the Hobbit \emph{bb-hosts(5)}
 configuration file. Default: \$BBHOME/etc/bb-hosts. 

 

\item[\index{BB}] Full path to the \emph{bb(1)}
 client program. Default: \$BBHOME/bin/bb. 

 

\item[\index{BBGEN}] Full path to the \emph{bbgen(1)}
 webpage generator program. Default: \$BBHOME/bin/bbgen. 

 


 


\end{description}

\subsection{URLS}
\begin{description}
\item[\index{BBSERVERWWWURL}] The root URL for the Hobbit webpages,
  without the hostname. This URL must be mapped to the ~/server/www/
  directory in your webserver configuration. See the sample Apache
  configuration in ~/server/etc/hobbit-apache.conf. 


 

\item[\index{BBSERVERCGIURL}] The root URL for the Hobbit CGI-scripts,
  without the hostname. This directory must be mapped to the
  ~/cgi-bin/ directory in your webserver configuration, and must be
  flagged as holding executable scripts. See the sample Apache
  configuration in ~/server/etc/hobbit-apache.conf. 


 

\item[\index{BBWEBHOST}] Initial part of the Hobbit URL, including
  just the protocol and the hostname,
  e.g. ``\url{http://www.foo.com}`` 


 

\item[\index{BBWEBHOSTURL}] Prefix for all of the static Hobbit
  webpages, e.g. ``\url{http://www.foo.com/hobbit}`` 


 

\item[\index{BBWEBHTMLLOGS}] URL prefix for the static HTML
  status-logs generated when BBLOGSTATUS=STATIC. Note that this
  setting is \textbf{discouraged} so this setting should not be used. 

 

\item[\index{BBWEB}] URL prefix (without hostname) of the Hobbit webpages. E.g. ``/hobbit''. 

 

\item[\index{BBSKIN}] URL prefix (without hostname) of the Hobbit graphics. E.g. ``/hobbit/gifs''. 

 

\item[\index{BBHELPSKIN}] URL prefix (without hostname) of the Hobbit on-line help files. E.g ``/hobbit/help''. 

 

\item[\index{BBMENUSKIN}] URL prefix (without hostname) of the Hobbit menu files. E.g ``/hobbit/menu''. 

 

\item[\index{BBNOTESSKIN}] URL prefix (without hostname) of the Hobbit on-line notes files. E.g ``/hobbit/notes''. 

 

\item[\index{BBREPURL}] URL prefix (without hostname) of the Hobbit availability reports. E.g. ``/hobbit/rep''. 

 

\item[\index{BBSNAPURL}] URL prefix (without hostname) of the Hobbit snapshots. E.g. ``/hobbit/snap''. 

 

\item[\index{BBWAP}] URL prefix (without hostname) of the Hobbit WAP/WML files. E.g. ``/hobbit/wml''. 

 

\item[\index{CGIBINURL}] URL prefix (without hostname) of the Hobbit CGI-scripts. Default: \$BBSERVERCGIURL . 

 

\item[\index{COLUMNDOCURL}] Format string used to build a link to the
  documentation for a column heading. Default:
  ``\$CGIBINURL/hobbitcolumn.sh?\%s'', which causes links to use the
  \emph{hobbitcolumn.sh(1)}

 script to document a column. 

 


 


\end{description}
\subsection{SETTINGS FOR SENDING MESSAGES TO HOBBIT}
\begin{description}
\item[\index{BBDISP}] The IP-address used to contact the \emph{hobbitd(8)}
 service. Used by clients and the tools that perform network tests. Default: \$BBSERVERIP 

 

\item[\index{BBDISPLAYS}] List of IP-adresses. Clients and network
  test tools will try to send status reports to a Hobbit server
  running on each of these adresses. This setting is only used if
  BBDISP=0.0.0.0. 


 

\item[\index{PAGELEVELS}] Compatibility setting for Big Brother: List
  of colors that are considered ``critical'' and therefore will
  trigger an alert. Not used by Hobbit. 


 

\item[\index{BBPAGE}] Compatibility setting for Big Brother: This is
  the IP-address of the server where a BBPAGER service is running. It
  is not used by Hobbit. 


 

\item[\index{BBPAGERS}] Compatibility setting for Big Brother: List of
  servers running the BBPAGER service, used if BBPAGE=0.0.0.0. It is
  not used by Hobbit. 


 

\item[\index{BBPORT}] The portnumber for used to contact the \emph{hobbitd(8)}
 service. Used by clients and the tools that perform network tests. Default: 1984. 

 

\item[\index{DOCOMBO}] Compatibility setting for Big Brother. Controls
  whether so send combo-messages or not. Ignored by Hobbit. 


\item[\index{BBMAXMSGSPERCOMBO}] The maximum number of status messages
  to combine into one combo message. You may need to lower this number
  of your BBDISPLAY server has trouble keeping up with the incoming
  status messages from bbtest-net. Default: 100. 


\item[\index{BBSLEEPBETWEENMSGS}] Length of a pause introduced between
  each successive transmission of a combo-message by bbtest-net. You
  may have to increase this value to give your BBDISPLAY server time
  to process one combo message before the next one arrives. This
  number defines how many microseconds to wait between the
  messages. Default: 0 (send messages as quickly as possible). 


\end{description}
\subsection{HOBBITD SETTINGS}


\begin{description}

\item[\index{ALERTCOLORS}] Comma-separated list of the colors that may
  trigger an alert-message. The default is ``red,yellow,purple''. Note
  that alerts may further be generated or suppresed based on the
  configuration in the \emph{hobbit-alerts.cfg(5)} file. 

 

\item[\index{OKCOLORS}] Comma-separated list of the colors that may
  trigger a recovery-message. The default is ``green,clear,blue''. 


 

\item[\index{ALERTREPEAT}] How often alerts get repeated while a
  status is in an alert state. This is the default setting, which may
  be changed in the \emph{hobbit-alerts.cfg(5)} file. 


 

\item[\index{BBGHOSTS}] Controls how status messages from unknown
  hosts (i.e. hosts not listed in the bb-hosts file) are handled. 


 \textbf{BBGHOSTS=1:} Causes the status report to be silently
 discarded. This is the default behaviour in Hobbit. 



 \textbf{BBGHOSTS=2:}
 Discards the status report, but keep track of the hostname and report
 it on the hobbitd status page. 



  When BBGHOSTS is set to 1 or 2, the hostnames in incoming
  status-messages is matched without any case-sensitivity, unlike
  normal Big Brother which is case-sensitive in hostnames. So with
  BBGHOSTS set to 1 or 2, ``WWW.FOO.COM'' and ``www.foo.com`` are
  considered to be the same host. If necessary, the incoming hostname
  will be changed to match the way it is written in the bb-hosts file,
  changing case as needed. 



 

\item[\index{MAXMSG\_STATUS}] The maximum size of a ``status'' message
  in kB, default: 256. Status messages are the ones that end up as
  columns on the web display. The default size should be adequate in
  most cases, but some extension scripts can generate very large
  status messages - close to 1024 kB. You should only change this if
  you see messages in the hobbitd log file about status messages being
  truncated. 


 

\item[\index{MAXMSG\_CLIENT}] The maximum size of a ``client'' message
  in kB, default: 512. ``client'' messages are generated by the Hobbit
  client, and often include large process-listings. You should only
  change this if you see messages in the hobbitd log file about client
  messages being truncated. 


 

\item[\index{MAXMSG\_DATA}] The maximum size of a ``data'' message in
  kB, default: 256. ``data'' messages are typically used for client
  reports of e.g. netstat or vmstat data. You should only change this
  setting if you see messages in the hobbitd log file about data
  messages being truncated. 


 

\item[\index{MAXMSG\_NOTES}] The maximum size of a ``notes'' message
  in kB, default: 256. ``notes'' messages provide a way for uploading
  documentation about a host to Hobbit; it is not enabled by
  default. If you want to upload large documents, you may need to
  change this setting. 


 

\item[\index{MAXMSG\_STACHG}] The maximum size of a ``status change''
  message in kB, default: Current value of the MAXMSG\_STATUS
  setting. Status-change messages occur when a status changes
  color. There is no reason to change this setting. 


 

\item[\index{MAXMSG\_PAGE}] The maximum size of a ``page'' message in
  kB, default: Current value of the MAXMSG\_STATUS setting. ``page''
  messages are alerts, and include the status message that triggers
  the alert. There is no reason to change this setting. 



\item[\index{MAXMSG\_ENADIS}] The maximum size of an ``enadis''
  message in kB, default: 32. ``enadis'' are small messages used when
  enabling or disabling hosts and tests, so the default size should be
  adequate. 


 

\item[\index{MAXMSG\_CLICHG}] The maximum size of a ``client change''
  message in kB, default: Current value of the MAXMSG\_CLIENT
  setting. Client-change messages occur when a status changes color to
  one of the alert-colors, usually red, yellow and purple. There is no
  reason to change this setting. 



\end{description}

\subsection{HOBBITD\_HISTORY SETTINGS}


 \begin{description}
\item[\index{BBALLHISTLOG}] If set to TRUE, \emph{hobbitd\_history(8)}
  will update the \$BBHIST/allevents file logging all changes to a
  status. The allevents file is used by the \emph{bb-eventlog.cgi(1)}
  tool to show the list of recent events on the BB2 webpage. 


 

\item[\index{BBHOSTHISTLOG}] If set to TRUE,
  \emph{hobbitd\_history(8)} will update the host-specific eventlog
  that keeps record of all status changes for a host. This logfile is
  not used by any Hobbit tool. 


 

\item[\index{SAVESTATUSLOG}] If set to TRUE, \emph{hobbitd\_history(8)}
 will save historical detailed status-logs to the \$BBHISTLOGS directory. 

 


 


\end{description}

\subsection{HOBBITD\_ALERT SETTINGS}


 \begin{description}
\item[\index{MAIL}] Command used to send alerts via e-mail, including a ``Subject:'' header in the mail. Default: ``mail -s'' 

 

\item[\index{MAILC}] Command used to send alerts via e-mail in a form that does not have a ``Subject'' in the mail. Default: ``mail'' 

 

\item[\index{SVCCODES}] Maps status-columns to numeric
  service-codes. The numeric codes are used when sending an alert
  using a script, where the numeric code of the service is provided in
  the BBSVCNUM variable. 


 


 


\end{description}

\subsection{HOBBITD\_RRD SETTINGS}


 \begin{description}

\item[\index{TEST2RRD}] List of ``COLUMNNAME[=RRDSERVICE]'' settings,
  that define which status- and data-messages have a corresponding RRD
  graph. You will normally not need to modify this, unless you have
  added a custom TCP-based test to the bb-services file, and want to
  collect data about the response-time, OR if you are using the
  \emph{hobbitd\_rrd(8)} external script mechanism to collect data
  from custom tests. Note: All TCP tests are automatically added. 



  This is also used by the \emph{bb-hostsvc.cgi(1) } script to
  determine if the detailed status view of a test should include a
  graph. 



 

\item[\index{GRAPHS}] List of the RRD databases, that should be shown
  as a graph on the ``trends'' column. 


 

\item[\index{NORRDDISKS}] This is used to disable the tracking of
  certain filesystems. By default all filesystems reported by a client
  are tracked. In some cases you may want to disable this for certain
  filesystems, e.g. database filesystems since they are always
  completely full. This setting is a regular expression that is
  matched against the filesystem name (the Unix mount-point, or the
  Windows disk-letter) - if the filesystem name matches this
  expression, then it will not be tracked by Hobbit.  

 Note: Setting this does not affect filesystems that are already being
 tracked by Hobbit - to remove them, you must remove the RRD files for
 the unwanted filesystems from the ~hobbit/data/rrd/HOSTNAME/
 directory. 


 

\item[\index{RRDDISKS}] This is used to enable tracking of only
  selected filesystems (see the NORRDDISKS setting above). By default
  all filesystems are being tracked, setting this changes that default
  so that only those filesystems that match this pattern will be
  tracked. 


 


 


\end{description}

\subsection{BBTEST-NET NETWORK TEST SETTINGS}


 \begin{description}

\item[\index{BBLOCATION}] If this variable is defined, then only the
  hosts that have been tagged with ``NET:\$BBLOCATION'' will be tested
  by the bbtest-net tool. 


\item[\index{CONNTEST}] If set to TRUE, the connectivity (ping) test will be performed. 

 

\item[\index{IPTEST\_2\_CLEAR\_ON\_FAILED\_CONN}] If set to TRUE, then
  failing network tests go CLEAR if the conn-test fails. 


\item[\index{NONETPAGE}] List of network services (separated with
  $<$space$>$) that should go yellow upon failure instead of red. 


 

\item[\index{BBROUTERTEXT}] When using the ``router'' or ``depends''
  tags for a host, a failure status will include text that an
  ``Intermediate router is down''. With todays network topologies, the
  router could be a switch or another network device; if you define
  this environment variable the word ``router'' will be replaced with
  whatever you put into the variable. So to inform the users that an
  intermediate switch or router is down, use BBROUTERTEXT=''switch or
  router''. This can also be set on a per-host basis using the
  ``DESCR:hosttype:description'' tag in the \emph{bb-hosts(5)} file. 


 

\item[\index{NETFAILTEXT}] When a network test fails, the status
  message reports ``SERVICENAME not OK''. The ``not OK'' message can
  be changed via this variable, e.g. you can change it to ``FAILED''
  or customize it as you like. 


 

\item[\index{FPING}] The command used to run the \emph{hobbitping(1)}
  tool for the connectivity test. (The name FPING is due to the fact
  that the ``fping'' utility was used until Hobbit version 4.2). This
  may include suid-root wrappers and hobbitping options. Default:
  ``hobbitping'' 


\item[\index{TRACEROUTE}] Defines the location of the ``traceroute''
  tool and any options needed to run it. traceroute it used by the
  connectivity test when the ping test fails; if requested via the
  ``trace'' tag, the TRACEROUTE command is executed to try to
  determine the point in the network that is causing the problem. By
  default the command executed is ``traceroute -n -q 2 -w 2 -m 15''
  (no DNS lookup, max. 2 probes, wait 2 seconds per hop, max 15
  hops). 


  If you have the \emph{mtr(8)} tool installed - available from
  \url{http://www.bitwizard.nl/mtr/} - I strongly recommend using this
  instead. The recommended setting for mtr is ``/usr/sbin/mtr -c 2 -n
  --report'' (the exact path to the mtr utility may be different on
  your system). Note that mtr needs to be installed suid-root on most
  systems. 



 

\item[\index{NTPDATE}] Defines the \emph{ntpdate(1)} program used for
  the ``ntp'' test. Default: ``ntpdate'' 


\item[\index{RPCINFO}] Defines the \emph{rpcinfo(8)}
 program used for ``rpc'' tests. Default: ``rpcinfo'' 

\end{description}

\subsection{BBGEN WEBPAGE GENERATOR SETTINGS}


 \begin{description}

\item[\index{HOBBITLOGO}] HTML code that is inserted on all standard
  headers. The default is to add the text ``Hobbit'' in the upper-left
  corner of the page, but you can easily replace this with e.g. a
  company logo. If you do, I suggest that you keep it at about 30-35
  pixels high, and 100-150 pixels wide. 


 

\item[\index{MKBBLOCAL}] The string ``Pages hosted locally'' that
  appears above all of the pages linked from the main Hobbit webpage. 


 

\item[\index{MKBBSUBLOCAL}] The string ``Subpages hosted locally''
  that appears above all of the sub-pages linked from pages below the
  main Hobbit webpage. 


 

\item[\index{MKBBREMOTE}] The string ``Remote status display'' that
  appears about the summary statuses displayed on the min Hobbit
  webpage. 


 

\item[\index{MKBBTITLE}] HTML tags designed to go in a $<$FONT$>$ tag,
  to choose the font for titles of the webpages. 


 

\item[\index{MKBBROWFONT}] HTML tags designed to go in a $<$FONT$>$
  tag, to choose the font for row headings (hostnames) on the
  webpages. 


 

\item[\index{MKBBCOLFONT}] HTML tags designed to go in a $<$FONT$>$
  tag, to chose the font for column headings (test names) on the
  webpages. 


 

\item[\index{MKBBACKFONT}] HTML tags designed to go in a $<$FONT$>$
  tag, to chose the font for the acknowledgement text displayed on the
  status-log HTML page for an acknowledged status. 


 

\item[\index{ACKUNTILMSG}] When displaying the detailed status of an
  acknowledged test, Hobbit will include the time that the acknowledge
  expires using the print-format defined in this setting. You can
  define the timeformat using the controls in your systems
  \emph{strftime(3)} routine, and add the text suitable for your
  setup. 


 

\item[\index{BBDATEFORMAT}] On webpages generated by bbgen, the
  default header includes the current date and time. Normally this
  looks like ``Tue Aug 24 21:59:47 2004''. The BBDATEFORMAT controls
  the format of this timestamp - you can define the format using the
  controls in the \emph{strftime(3)} routine. E.g. to have it show up
  as ``2004-08-24 21:59:47 +0200'' you would set
  BBDATEFORMAT=''\%Y-\%m-\%d \%H:\%M:\%S \%z'' 


 

\item[\index{HOLIDAYFORMAT}] How holiday dates are displayed. The
  default is ``\%d/\%m'' which show the day and month. American users
  may want to change this to ``\%m/\%d'' to suit their preferred
  date-display style. This is a formatting string for the system
  \emph{strftime(3)} routine, so any controls available for this
  routine may be used. 


 

\item[\index{MKBB2COLREPEAT}] Inspired by Jeff Stoner's
  col\_repeat\_patch.tgz patch, this defines the maximum number of
  rows before repeating the column headings on a webpage. This sets
  the default value for the \emph{bbgen(1)} ``--maxrows'' option; if
  the command-line option is also specifed, then it overrides this
  environment variable. Note that unlike Jeff's patch, bbgen
  implements this for both the bb2.html page and all other pages
  (bb.html, subpages, bbnk.html). 


 

\item[\index{SUMMARY\_SET\_BKG}] If set to TRUE, then summaries will
  affect the color of the main Hobbit webpage. Default: FALSE. 


 

\item[\index{DOTHEIGHT}] The height (in pixels) of the icons showing
  the color of a status. Default: 16, which matches the default
  icons. 


 

\item[\index{DOTWIDTH}] The width (in pixels) of the icons showing the
  color of a status. Default: 16, which matches the default icons. 


 

\item[\index{CLIENTSVCS}] List of the status logs fed by data from the
  Hobbit client. These status logs will - if there are Hobbit client
  data available for the host - include a link to the raw data sent by
  the client. Default: cpu,disk,memory,procs,svcs. 


\item[\index{BBRSSTITLE}] If defined, this is the title of the RSS/RDF documents generated when \emph{bbgen(1)}
 is invoked with the ``--rss'' option. The default value is ``Hobbit Alerts''. 

 

\item[\index{WMLMAXCHARS}] Maximum size of a WAP/WML output ``card'' when generating these. Default: 1500. 

 

\item[\index{BBMKBB2EXT}] List of scripts to run as extensions to the
  BB2 page. Note that two scripts, ``eventlog.sh'' and ``acklog.sh''
  are handled specially: They are handled internally by bbgen, but the
  script names must be listed in this variable for this function to be
  enabled. 


 

\item[\index{BBHISTEXT}] List of scripts to run as extensions to a history page. 

 

\item[\index{BBREPWARN}] Default threshold for listing the
  availability as ``critical'' (red) when generating the availability
  report. This can be set on a per-host basis with the WARNPCT setting
  in \emph{bb-hosts(5).}

 Default: 97 (percent) 

 

\item[\index{BBGENREPOPTS}] Default bbgen options used for
  reports. This will typically include such options as
  ``--subpagecolumns'', and also ``--ignorecolumns'' if you wish to
  exclude certain tests from reports by default. 


 

\item[\index{BBGENSNAPOPTS}] Default bbgen options used by
  snapshots. This should be identical to the options you normally used
  when building Hobbit webpages. 



\end{description}

\subsection{FILES}
\textbf{~hobbit/server/etc/hobbitserver.cfg}


 
\subsection{SEE ALSO}
hobbit(7) 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{bb-hosts - Main Hobbit configuration file}
 
\subsection{SYNOPSIS}
\begin{description}
\item[\index{bb-hosts}]


\end{description}
\subsection{DESCRIPTION}
 The \emph{bb-hosts(5)}
 file is the most important configuration file for all of the Hobbit
 programs. This file contains the full list of all the systems
 monitored by Hobbit, including the set of tests and other
 configuration items stored for each host. 
 
\subsection{FILE FORMAT}
 Each line of the file defines a host. Blank lines and lines starting
 with a hash mark (\#) are treated as comments and ignored. Long lines
 can be broken up by putting a backslash at the end of the line and
 continuing the entry on the next line. 


  The format of an entry in the bb-hosts file is as follows:  
 
IP-addresshostname\#tag1tag2... 


  The IP-address and hostname are mandatory; all of the tags are
  optional. Listing a host with only IP-address and hostname will
  cause a network test to be executed for the host - the connectivity
  test is enabled by default, but no other tests. 



  The optional tags are then used to define which tests are relevant
  for the host, and also to set e.g. the time-interval used for
  availability reporting by \emph{bbgen(1)}




  An example of setting up the bb-hosts file is in the Hobbit on-line
  documentation (from the Help menu, choose ``Configuring
  Monitoring''). The following describes the possible settings in a
  bb-hosts file supported by Hobbit. 


 
\subsection{TAGS RECOGNIZED BY ALL TOOLS}


 \begin{description}
\item[\index{include filename}] This tag is used to include another file into
  the bb-hosts file at run-time, allowing for a large bb-hosts file to
  be split up into more manageable pieces. 


  The ``filename'' argument should point to a file that uses the same
  syntax as bb-hosts. The filename can be an absolute filename (if it
  begins with a '/'), or a relative filename - relative filenames are
  prefixed with the directory where the main bb-hosts file is located
  (usually \$BBHOME/etc/). 



  You can nest include tags, i.e. a file that is included from the
  main bb-hosts file can itself include other files. 



 

\item[\index{dispinclude filename}] Acts like the ``include'' tag, but only on
  the BBDISPLAY server. Can be used e.g. to put a group of hosts on
  multiple subpages, without having to repeat the host definitions. 


 

\item[\index{netinclude filename}] Acts like the ``include'' tag, but only on the BBNET server. 

 


 


\end{description}

\subsection{GENERAL PER-HOST OPTIONS }
\begin{description}
\item[\index{noclear}] Controls whether stale status messages go purple or
  clear when a host is down. Normally, when a host is down the client
  statuses (``cpu'', ``disk'', ``memory'' etc) will stop updating -
  this would usually make them go ``purple'' which can trigger
  alerts. To avoid that, Hobbit checks if the ``conn'' test has
  failed, and if that is true then the other tests will go ``clear''
  instead of purple so you only get alerts for the ``conn'' test. If
  you do want the stale statuses to go purple, you can use the
  ``noclear'' tag to override this behaviour. 


  Note that ``noclear'' also affects the behaviour of network tests; see below. 


 

\item[\index{prefer}] When a single host is defined multiple time in the
  bb-hosts file, bbgen tries to guess which definition is the best to
  use for the information used on the ``info'' column, or for the
  NOPROPRED and other bbgen-specific settings. Host definitions that
  have a ``noconn'' tag or an IP of 0.0.0.0 get lower priority. 


  By using the ``prefer'' tag you tell bbgen that this host definition should be used. 


  Note: This only applies to hosts that are defined multiple times in
  the bb-hosts file, although it will not hurt to add it on other
  hosts as well. 



 
\end{description}
\subsection{BBGEN DISPLAY OPTIONS}
 These tags are processed by the \emph{bbgen(1)}
 tool when generating the Hobbit webpages or reports. 

 \begin{description}
\item[page NAME [Page-title]] This defines a page at the level below
  the entry page. All hosts following the ``page'' directive appear on
  this page, until a new ``page'', ``subpage'' or ``subparent'' line
  is found. 


 

\item[subpage NAME [Page-title]] This defines a subpage in the second
  level below the entry page. You must have a previous ``page'' line
  to hook this subpage to. 


 

\item[subparent parentpage newpage [Page-title]] This is used to
  define subpages in whatever levels you may wish. Just like the
  standard ``subpage'' tag, ``subparent'' defines a new Hobbit
  webpage; however with ``subparent'' you explicitly list which page
  it should go as a subpage to. You can pick any page as the parent -
  pages, subpages or even other subparent pages. So this allows you to
  define any tree structure of pages that you like. 


  E.g. with this in bb-hosts: 


  
pageUSAUnitedStates  
subpageNYNewYork  
subparentNYmanhattanManhattandatacenters  
subparentmanhattanwallstreetWallStreetcenter 


  you get this hierarchy of pages: 


  
USA(UnitedStates)  
NY(NewYork)  
manhattan(Manhattandatacenters)  
wallstreet(WallStreetcenter) 


  Note: The parent page must be defined before you define the
  subparent. If not, the page will not be generated, and you get a
  message in the log file. 



  Note: bbgen is case-sensitive, when trying to match the name of the parent page. 


  The inspiration for this came from Craig Cook's mkbb.pl script, and
  I am grateful to Craig for suggesting that I implement it in
  bbgen. The idea to explicitly list the parent page in the
  ``subparent'' tag was what made it easy to implement. 



 

\item[group [group-title]]
\item[group-compress [group-title]] Defines a group of hosts, that
  appear together on the webpage, with a single header-line listing
  all of the columns. Hosts following the ``group'' line appear inside
  the group, until a new ``group'' or page-line is found. The two
  group-directives are handled identically by Hobbit and bbgen, but
  both forms are allowed for backwards compatibility. 


 

\item[group-sorted [group-title]] Same as the ``group'' line, but will
  sort the hosts inside the group so they appear in strict
  lexicographic order. 


 

\item[group-only COLUMN1|COLUMN2|COLUMN3 [group-title]] Same as the
  ``group'' and ``group-compress'' lines, but includes only the
  columns explicitly listed in the group. Any columns not listed will
  be ignored for these hosts. 


 

\item[group-except COLUMN1|COLUMN2|COLUMN3 [group-title]] Same as the
  ``group-only'' lines, but includes all columns EXCEPT those
  explicitly listed in the group. Any columns listed will be ignored
  for these hosts - all other columns are shown. 


 

\item[title Page, group or host title text] The ``title'' tag is used
  to put custom headings into the pages generated by bbgen, in front
  of page/subpage links, groups or hosts. 


  The title tag operates on the next item in the bb-hosts file following the title tag. 


  If a title tag precedes a host entry, the title is shown just before
  the host is listed on the status page. The column headings present
  for the host will be repeated just after the heading. 



  If a title tag precedes a group entry, the title is show just before the group on the status page. 


  If a title tag precedes a page/subpage/subparent entry, the title
  text replaces the normal ``Pages hosted locally'' heading normally
  inserted by Hobbit. This appears on the page that links to the
  subpages, not on the subpage itself. To get a custom heading on the
  subpage, you may want to use the ``--pagetext-heading'' when running
  \emph{bbgen(1)}




 

\item[NAME:hostname] Overrides the default hostname used on the
  overview web pages. If ``hostname'' contains spaces, it must be
  enclosed in double quotes, e.g. NAME:''R\&D Oracle Server'' 


 

\item[CLIENT:hostname] Defines an alias for a host, which will be used
  when identifying status messages. This is typically used to
  accomodate a local client that sends in status reports with a
  different hostname, e.g. if you use hostnames with domains in your
  Hobbit configuration, but the client is a silly Window box that does
  not include the hostname. Or vice versa. Whatever the reason, this
  can be used to match status reports with the hosts you define in
  your bb-hosts file. It causes incoming status reports with the
  specified hostname to be filed using the hostname defined in
  bb-hosts. 


 

\item[NOCOLUMNS:column[,column]] Used to drop certain of the status
  columns generated by the Hobbit client. \textbf{column}

 is one of \textbf{cpu}
, \textbf{disk}
, \textbf{files}
, \textbf{memory}
, \textbf{msgs}
, \textbf{ports}
, \textbf{procs}
. This setting stops these columns from being updated for the
 host. Note: If the columns already exist, you must use the
 \emph{bb(1)} utility to \textbf{drop} them, or they will go purple. 


 

\item[COMMENT:Host comment] Adds a small text after the hostname on
  the webpage. This can be used to describe the host, without
  completely changing its display-name as the NAME: tag does. If the
  comment includes whitespace, it must be in double-quotes,
  e.g. COMMENT:''Sun webserver'' 


 

\item[DESCR:Hosttype:Description] Define some informational text about
  the host. The ``Hosttype'' is a text describing the type of this
  device - ``router'', ``switch'', ``hub'', ``server'' etc. The
  ``Description'' is an informational text that will be shown on the
  ``Info'' column page; this can e.g. be used to store information
  about the physical location of the device, contact persons etc. If
  the text contain whitespace, you must enclose it in double-quotes,
  e.g. DESCR:''switch:4th floor Marketing switch'' 


 \item[CLASS:Classname] Force the host to belong to a specific
  class. Class-names are used when configuring log-file monitoring
  (they can be used as references in \emph{client-local.cfg(5)} and
  \emph{hobbit-clients.cfg(5)} to group logfile checks). Normally,
  class-names are controlled on the client by starting the Hobbit
  client with the ``--class=Classname'' option. If you specify it in
  the bb-hosts file on the Hobbit server, it overrides any classname
  that the client reports. 


 

\item[\index{dialup}] The keyword ``dialup'' for a host means that it is OK
  for it to be off-line - this should not trigger an alert. All
  network tests will go ``clear'' upon failure, and any missing
  reports from e.g. cpu- and disk-status will not go purple when they
  are not updated. 


 

\item[\index{nobb2}] Ignore this host on the BB2 page. Even if it has an
  active alert, it will not be included in the BB2 page. This also
  removes the host from the event-log display. 


 

\item[\index{nodisp}] Ignore this host completely when generating the Hobbit
  webpages. Can be useful for monitoring a host without having it show
  up on the webpages, e.g. because it is not yet in production use. Or
  for hiding a host that is shown only on a second pageset. 


 

\item[TRENDS:[*,][![graph,...]]] Defines the RRD graphs to include in
  the ``trends'' column generated by bbgen. This option syntax is
  complex.  

 If this option is not present, bbgen provides graphs matching the
 standard set of RRD files: la, disk, memory, users, vmstat, iostat,
 netstat, tcp, bind, apache, sendmail  

 If this option is specified, the list of graphs to include start out as being empty (no graphs).  
 To include all default graphs, use an asterisk. E.g. ``TRENDS:*''  
 To exclude a certain graph, speficy it prefixed with '!'. E.g. to see
 all graphs except users: ``TRENDS:*,!users''  

 The netstat, vmstat and tcp graphs have many ``subgraphs''. Which of these are shown can be speficied like this:
 ``TRENDS:*,netstat:netstat2|netstat3,tcp:http|smtp|conn'' This will
 show all graphs, but instead of the normal netstat graph, there will
 be two: The netstat2 and netstat3 graphs. 

Instead of the combined tcp graphs showing all services, there will be three: One
 for each of the http, conn and smtp services. 


 

\item[DOWNTIME=[columns:]day:starttime:endtime:cause[,day:starttime:endtime:cause]]
  This tag can be used to ignore failed checks during specific times
  of the day - e.g. if you run services that are only monitored
  e.g. Mon-Fri 8am-5pm, or you always reboot a server every Monday
  between 5 and 6 pm. 


  The timespecification consists of 


 \textbf{day:}
 \textbf{W}
 means Mon-Fri (``weekdays''), \textbf{*}
 means all days, \textbf{0}
 .. \textbf{6}
 = Sunday .. Saturday. Listing multiple days is possible, e.g. ``60'' is valid meaning ``Saturday and Sunday''. 


 \textbf{starttime:}
 Time to start showing errors, must be in 24-hour clock format as HHMM
 hours/minutes. E.g. for 8 am enter ``0800'', for 9.30 pm enter
 ``2130'' 



 \textbf{endtime:}

 Time to stop showing errors. 


 If necessary, multiple periods can be specified. E.g. to monitor a
 site 24x7, except between noon and 1 pm, use
 DOWNTIME=*:0000:1159,*:1300:2359 



 The interval between starttime and endtime may cross midnight, e.g. \textbf{*:2330:0200}
 would be valid and have the same effect as \textbf{*:2330:2400,*:0000:0200}. 


  What happens is that if a test fails during the specified time, it
  is reported with status BLUE instead of yellow or red. Thus you can
  still see when the service was unavailable, but alarms will not be
  triggered and the downtime is not counted in the availability
  calculations generated by the Hobbit reports. 



  The ``columns'' setting is optional - it may be a comma-separated
  list of status columns in which case the DOWNTIME setting only
  applies to these columns. 



  The ``cause'' string (optional) is a text that will be displayed on
  the status web page to explain thy the system is down. 



 

\item[SLA=day:starttime:endtime[,day:starttime:endtime]] This tag is now deprecated. Use the DOWNTIME tag instead. 

  This tag works the opposite of the DOWNTIME tag - you use it to
  specify the periods of the day that the service should be
  green. Failures OUTSIDE the SLA interval are reported as blue. 


\item[holidays=HOLIDAYSETNAME] Define which set of holidays to use for
  a host. Holiday sets are defined in the
  \emph{hobbit-holidays.cfg(5)} configuration file. If omitted, the
  set given in the HOLIDAYS setting in the \emph{hobbitserver.cfg(5)}
  file is used. 

\end{description}

\subsection{HOBBIT TAGS FOR THE CRITICAL SYSTEMS OVERVIEW PAGE}
\textbf{NOTE:}
 The ``NK'' set of tags is deprecated. They will be supported for
 Hobbit 4.x, but will be dropped in version 5. It is recommended that
 you move your critical systems view to the
 \emph{hobbit-nkview.cgi(1)} viewer, which has a separate
 configuration tool, \emph{hobbit-nkedit.cgi(1)} with more facilities
 than the NK tags in bb-hosts. 


 bbgen will create three sets of pages: The main page bb.html, the
 all-non-green-statuses page (bb2.html), and a specially reduced
 version of bb2.html with only selected tests (bbnk.html). This page
 includes selected tests that currently have a red or yellow status. 



 \begin{description}
\item[NK:testname[,testname]] Define the tests that you want included
  on the bbnk page. E.g. if you have a host where you only want to see
  the http tests on bbnk.html, you specify it as 


  
12.34.56.78www.acme.com\#\url{http://www.acme.com/}NK:http 


  If you want multiple tests for a host to show up on the bbnk.html
  page, specify all the tests separated by commas. The test names
  correspond to the column names (e.g. https tests are covered by an
  ``NK:http'' tag). 



\item[NKTIME=day:starttime:endtime[,day:starttime:endtime]] This tag limits the time when an active alert is presented on the NK webpage. 

  By default, tests with a red or yellow status that are listed in the
  ``NK:testname'' tag will appear on the NK page. However, you may not
  want the test to be shown outside of normal working hours - if, for
  example, the host is not being serviced during week-ends. 



  You can then use the NKTIME tag to define the time periods where the
  alert will show up on the NK page. The NKTIME syntax is identical to
  the one used for DOWNTIME specification. 


\end{description}

\subsection{HOBBIT TAGS FOR THE WML (WAP) CARDS}
 If bbgen is run with the ``--wml'' option, it will generate a set of
 WAP-format output ``cards'' that can be viewed with a WAP-capable
 device, e.g. a PDA or cell-phone. 


 \begin{description}
\item[WML:[+|-]testname[,[+|-]testname]] This tag determines which
  tests for this hosts are included in the WML (WAP) page. Syntax is
  identical to the NK: tag. 


  The default set of WML tests are taken from the --wml commandline
  option. If no ``WML:'' tag is specified, the ``NK:'' tag is used if
  present. 




\end{description}

\subsection{HOBBIT STATUS PROPAGATION OPTIONS}
 These tags affect how a status propagates upwards from a single test
 to the page and higher. This can also be done with the command-line
 options --nopropyellow and --nopropred, but the tags apply to
 individual hosts, whereas the command line options are global. 


 \begin{description}
\item[NOPROPRED:[+|-]testname[,[+|-]testname]] This tag is used to
  inhibit a yellow or red status from propagating upwards - i.e. from
  a test status color to the (sub)page status color, and further on to
  bb.html or bb2.html 


  If a host-specific tag begins with a '-' or a '+', the host-specific
  tags are removed/added to the default setting from the command-line
  option. If the host-specific tag does not begin with a '+' or a '-',
  the default setting is ignored for this host and the NOPROPRED
  applies to the tests given with this tag. 



  E.g.: bbgen runs with
  ``--nopropred=ftp,smtp''. ``NOPROPRED:+dns,-smtp'' gives a NOPROPRED
  setting of ``ftp,dns'' (dns is added to the default, ftp is
  removed). ``NOPROPRED:dns'' gives a setting of ``dns'' only (the
  default is ignored). 



  Note: If you set use the ``--nopropred=*'' commandline option to
  disable propagation of all alerts, you cannot use the ``+'' and
  ``-'' methods to add or remove from the wildcard setting. In that
  case, do not use the ``+'' or ``-'' setting, but simply list the
  required tests that you want to keep from propagating. 


\item[NOPROPYELLOW:[+|-]testname[,[+|-]testname]] Similar to
  NOPROPRED: tag, but applies to propagating a yellow status upwards. 


\item[NOPROPPURPLE:[+|-]testname[,[+|-]testname]] Similar to
  NOPROPRED: tag, but applies to propagating a purple status upwards. 


 

\item[NOPROPACK:[+|-]testname[,[+|-]testname]] Similar to NOPROPRED:
  tag, but applies to propagating an acknowledged status upwards. 


\end{description}

\subsection{HOBBIT AVAILABILITY REPORT OPTIONS}
 These options affect the way the Hobbit availability reports are processed (see \emph{bb-rep.cgi(1)}
 for details about availability reports). 

 \begin{description}
\item[REPORTTIME=day:starttime:endtime[,day:starttime:endtime]] This
  tag defines the time interval where you measure uptime of a service
  for reporting purposes. 


  When bbgen generates a report, it computes the availability of each
  service - i.e. the percentage of time that the service is reported
  as available (meaning: not red). 



  By default, this calculation is done on a 24x7 basis, so no matter
  when an outage occurs, it counts as downtime. 



  The REPORTTIME tag allows you to specify a period of time other than
  24x7 for the service availability calculation. If you have systems
  where you only guarantee availability from e.g. 7 AM to 8 PM on
  weekdays, you can use  

 
REPORTTIME=W:0700:2000  
 and the availability calculation will only be performed for the
 service with measurements from this time interval. 



  The syntax for REPORTTIME is the same as the one used by the DOWNTIME parameter. 


  When REPORTTIME is specified, the availability calculation happens like this: 


  * Only measurements done during the given time period is used for the calculation.  
 * ``blue'' time reduces the length of the report interval, so if you
 are generating a report for a 10-hour period and there are 20 minutes
 of ``blue'' time, then the availability calculation will consider the
 reporting period to be 580 minutes (10 hours minus 20 minutes). This
 allows you to have scheduled downtime during the REPORTTIME interval
 without hurting your availability; this is (I believe) the whole idea
 of the downtime being ``planned''.  

 * ``red'' and ``clear'' status counts as downtime; ``yellow'' and
   ``green'' count as uptime. ``purple'' time is ignored. 



  The availability calculation correctly handles status changes that cross into/out of a REPORTTIME interval. 


  If no REPORTTIME is given, the standard 24x7 calculation is used. 


 

\item[WARNPCT:percentage] BB's reporting facility uses a computed
  availability threshold to color services green (100\% available),
  yellow (above threshold, but less than 100\%), or red (below
  threshold) in the reports. 


  This option allows you to set the threshold value on a host-by-host
  basis, instead of using a global setting for all hosts. The
  threshold is defined as the percentage of the time that the host
  must be available, e.g. ``WARNPCT:98.5'' if you want the threshold
  to be at 98.5\% 



 

\end{description}

\subsection{NETWORK TEST SETTINGS}
\begin{description}
\item[\index{testip}] By default, Hobbit will perform a name lookup of the
  hostname to get the IP address it will use for network tests. This
  tag causes Hobbit to use the IP listed in the bb-hosts file. 


 

\item[NET:location] This tag defines the host as being tested from a
  specific location. If bbtest-net sees that the environment variable
  BBLOCATION is set, it will only test the hosts that have a matching
  ``NET:location'' tag in the bb-hosts file. So this tag is useful if
  you have more than one BBNET system, but you still want to keep a
  consolidated bb-hosts file for all your systems. 


  Note: The ``--test-untagged'' option modifies this behaviour, see \emph{bbtest-net(1)}



 

\item[\index{noclear}] Some network tests depend on others. E.g. if the host
  does not respond to ping, then there's a good chance that the entire
  host is down and all network tests will fail. Or if the http server
  is down, then any web content checks are also likely to fail. To
  avoid floods of alerts, the default behaviour is for bbtest-net to
  change the status of these tests that fail because of another
  problem to ``clear'' instead of ``red''. The ``noclear'' tag
  disables this behaviour and causes all failing tests to be reported
  with their true color. 


  This behaviour can also be implemented on a per-test basis by putting the ``~`` flag on any network test. 


  Note that ``noclear'' also affects whether stale status messages
  from e.g. a client on the host go purple or clear when the host is
  down; see the ``noclear'' description in the ``GENERAL PER-HOST
  OPTIONS'' section above. 

\item[\index{nosslcert}] Disables the standard check of any SSL certificates
  for this host. By default, if an SSL-enabled service is tested, a
  second test result is generated with information about the SSL
  certificate - this tag disables the SSL certificate checks for the
  host. 


 

\item[ssldays=WARNDAYS:ALARMDAYS] Define the number of days before an
  SSL certificate expires, in which the sslcert status shows a warning
  (yellow) or alarm (red) status. These default to the values from the
  ``--sslwarndays'' and ``--sslalarmdays'' options for the
  \emph{bbtest-net(1)} tool; the values specified in the ``ssldays''
  tag overrides the default. 


 

\item[depends=(testA:host1/test1,host2/test2),(testB:host3/test3),[...]]
  This tag allows you to define dependencies betweeen tests. If
  ``testA'' for the current host depends on ``test1'' for host
  ``host1'' and test ``test2'' for ``host2'', this can be defined with 


  
depends=(testA:host1/test1,host2/test2) 


  When deciding the color to report for testA, if either host1/test1
  failed or host2/test2 failed, if testA has failed also then the
  color of testA will be ``clear'' instead of red or yellow. 



  Since all tests are actually run before the dependencies are
  evaluated, you can use any host/test in the dependency - regardless
  of the actual sequence that the hosts are listed, or the tests
  run. It is also valid to use tests from the same host that the
  dependency is for. E.g. 



  
1.2.3.4foo\#\url{http://foo/}webmindepends=(webmin:foo/http) 


  is valid; if both the http and the webmin tests fail, then webmin will be reported as clear. 


  Note: The ``depends'' tag is evaluated on the BBNET server while
  running the network tests. It can therefore only refer to other
  network tests that are handled by the same BBNET server - there is
  currently no way to use the e.g. the status of locally run tests
  (disk, cpu, msgs) or network tests from other BBNET servers in a
  dependency definition. Such dependencies are silently ignored. 



\item[badTEST[-weekdays-starttime-endtime]:x:y:z] Normally when a
  network test fails, the status changes to red immediately. With a
  ``badTEST:x:y:z'' tag this behaviour changes:  

 * While ``z'' or more successive tests fail, the column goes RED.  
 * While ``y'' or more successive tests fail, but fewer than ``z'', the column goes YELLOW.  
 * While ``x'' or more successive tests fail, but fewer than ``y'', the column goes CLEAR.  
 * While fewer than ``x'' successive tests fail, the column stays GREEN. 

  The optional timespecification can be used to limit this ``badTEST''
  setting to a particular time of day, e.g. to require a longer period
  of downtime before raising an alarm during out-of-office hours. The
  time-specification uses:  

 * Weekdays: The weekdays this badTEST tag applies, from 0 (Sunday)
   through 6 (Saturday). Putting ``W'' here counts as ``12345'',
   i.e. all working days. Putting ``*'' here counts as all days of the
   week, equivalent to ``0123456''.  

 * starttime and endtime are specified using 24-hour clocks,
   e.g. ``badTEST-W-0900-2000'' is valid for working days between 9 AM
   (09:00) and 8 PM (20:00). 



  When using multiple badTEST tags, the LAST one specified with a matching time-spec is used. 


  Note: The ``TEST'' is replaced by the name of the test, e.g. 


  
12.34.56.78www.foo.com\#\url{http://www.foo.com/}badhttp:1:2:4 


  defines a http test that goes ``clear'' after the first failure,
  ``yellow'' after two successive failures, and ``red'' after four
  successive failures. 



  For the other network tests, use ``badftp'', ``badssh'' etc. 


 



\end{description}
\subsection{CONNECTIVITY (PING) TEST}
 These tags affect the behaviour of the bbtest-net connectivity test. 

 \begin{description}
\item[\index{noping}] Disables the ping-test, but will keep the ``conn'' column on the web display with a notice that it has been disabled. 

 

\item[\index{noconn}] Disables the ping-test, and does not put a ``conn'' column on the web display. 

 

\item[\index{conn}] The ``conn'' test (which does a ping of the host) is
  enabled for all hosts by default, and normally you just want to
  disable it using ``noconn'' or ``noping''. However, on the rare
  occasion where you may want to check that a host is NOT up, you can
  specify it as an explicit test, and use the normal test modifiers,
  e.g. ``!conn'' will be green when the host is NOT up, and red if it
  does appear on the network. 


  The actual name of the tag - ``conn'' by default - depends on the ``--ping=TESTNAME'' option for bbtest-net, as that decides the testname for the connectivity test. 


 

\item [conn=best,|worst, IP1[,IP2...]] This adds additional
  IP-adresses that are pinged during the normal ``conn'' test. So the
  normal ``conn'' test must be enabled (the default) before this tag
  has any effect. The IP-adresses listed here are pinged in addition
  to the main IP-address. 


  When multiple IP's are pinged, you can choose if ALL IP's must
  respond (the ``worst'' method), or AT LEAST one IP must respond (the
  ``best'' setting). All of the IP's are reported in a single ``conn''
  status, whose color is determined from the result of pinging the
  IP's and the best/worst setting. The default method is ``best'' - so
  it will report green if just one of the IP's respond to ping. 



 

\item[badconn[-weekdays-starttime-endtime]:x:y:z] This is taken
  directly from the ``fping.sh'' connectivity- testing script, and is
  used by bbtest-net when it runs with ping testing enabled (the
  default). See the description of the ``badTEST'' tag. 


 

\item[route:router1,router2,....] This tag is taken from the ``fping.sh'' script, and is used by bbtest-net when run with the ``--ping'' option to enable ping testing. 

  The router1,router2,... is a comma-separated list of hosts elsewhere in the bb-hosts file. You cannot have any spaces in the list - separate hosts with commas. 


  This tag changes the color reported for a ping check that fails,
  when one or more of the hosts in the ``route'' list is also down. A
  ``red'' status becomes ``yellow'' - other colors are unchanged. The
  status message will include information about the hosts in the
  router-list that are down, to aid tracking down which router is the
  root cause of the problem. 



  Note: Internally, the ping test will still be handled as ``failed'',
  and therefore any other tests run for this host will report a status
  of ``clear''. 



 

\item[route\_LOCATION:router1,router2,...] If the BBLOCATION
  environment variable is defined, a tag of ``route\_BBLOCATION:'' is
  recognized by bbtest-net with the same effect as the normal
  ``route:'' tag (see above). This allows you to have different route:
  tags for each BBNET server. The actual text for the tag then must
  match the value you have for the BBLOCATION setting. E.g. with
  BBLOCATION=dmz, the tag becomes ``route\_dmz:'' 


 

\item[\index{trace}] If the connectivity test fails, run a ``traceroute'' and
  include the output from this in the status message from the failed
  connectivity test. Note: For this to work, you may have to define
  the TRACEROUTE environment variable, see \emph{hobbitserver.cfg(5)}



 

\item[\index{notrace}] Similar to the ``trace'' option, this disables the
  running of a traceroute for the host after a failed connectivity
  test. It is only used if running traceroute is made the default via
  the --trace option. 


 


\end{description}

\subsection{SIMPLE NETWORK TESTS}
 These tests perform a simple network test of a service by connecting
 to the port and possibly checking that a banner is shown by the
 server. 


  How these tests operate are configured in the \emph{bb-services(5)}
  configuration file, which controls which port to use for the
  service, whether to send any data to the service, whether to check
  for a response from the service etc. 



  You can modify the behaviour of these tests on a per-test basis by
  adding one or more modifiers to the test: \textbf{:NUMBER} changes
  the port number from the default to the one you specify for this
  test. E.g. to test ssh running on port 8022, specify the test as
  \textbf{ssh:8022}. 


 \textbf{:s}
 makes the test silent, i.e. it does not send any data to the
 service. E.g. to do a silent test of an smtp server, enter
 \textbf{smtp:s}. 



  You can combine these two: \textbf{\url{ftp:8021:s}}
 is valid. 


  If you must test a service from a multi-homed host (i.e. using a
  specific source IP-address instead of the one your operating system
  provides), you can use the modifier ``@IPADDRESS'' at the end of the
  test specification, \textbf{after} any other modifiers or port
  number. ``IPADDRESS'' must be a valid dotted IP-address (not
  hostname) which is assigned to the host running the network tests. 



  The name of the test also determines the columnname that the test
  result will appear with in the Hobbit webpages. 



  By prefixing a test with ``!'' it becomes a reverse test: Hobbit
  will expect the service NOT to be available, and send a green status
  if it does NOT respond. If a connection to the service succeeds, the
  status will go red. 



  By prefixing a test with ``?'' errors will be reported with a
  ``clear'' status instead of red. This is known as a test for a
  ``dialup'' service, and allows you to run tests of hosts that are
  not always online, without getting alarms while they are off-line. 



 \begin{description}
\item[ftp ssh telnet smtp pop3 imap nntp rsync clamd oratns qmtp qmqp]
  These tags are for testing services offering the FTP, Secure Shell
  (ssh), SMTP, POP3, IMAP, NNTP, rsync, CLAM antivirus daemon (clamd),
  Oracle TNS listener (oratns), qmail QMTP and QMQP protocols. 


 

\item[ftps telnets smtps pop3s imaps nntps] These tags are for testing
  of the SSL-tunneled versions of the standard ftp, telnet, smtp,
  pop3, imap and nntp protocols. If Hobbit was configured with support
  for SSL, you can test these services like any other network service
  - bbtest-net will setup an SSL-encrypted session while testing the
  service. The server certificate is validated and information about
  it sent in the ``sslcert'' column. Note that smtps does not have a
  standard portnumber assignment, so you will need to enter this into
  the bb-services file or your /etc/services file. 


 

\item[\index{bbd}] Test that a Big Brother compatible daemon is running. This
  check works both for the Hobbit \emph{hobbitd(8)} daemon, and the
  original Big Brother bbd daemon. 


 


 


\end{description}

\subsection{DNS SERVER TESTS}
 These tags are used to setup monitoring of DNS servers. 

 \begin{description}
\item[\index{dns}] Simple DNS test. It will attempt to lookup the A record for the hostname of the DNS server. 

 

\item[\index{dig}] This is an alias for the ``dns'' test. In bbtest-net, the
  ``dns'' and ``dig'' tests are handled identically, so all of the
  facilities for testing described for the ``dns'' test are also
  available for the ``dig'' test. 


 

\item[dns=hostname]
\item[dns=TYPE:lookup[,TYPE:lookup...]] The default DNS tests will
  attempt a DNS lookup of the DNS' servers own hostname. You can
  specify the hostname to lookup on a DNS server by listing it on each
  test. 


  The second form of the test allows you to perform multiple queries
  of the DNS server, requesting different types of DNS records. The
  TYPE defines the type of DNS data: A (IP-address), MX (Mail
  eXchanger), PTR (reverse), CNAME (alias), SOA (Start-Of-Authority),
  NS (Name Server) are among the more common ones used. The ``lookup''
  is the query. E.g. to lookup the MX records for the ``foo.com''
  domain, you would use ``dns=mx:foo.com''. Or to lookup the
  nameservers for the ``bar.org'' domain, ``dns=ns:bar.org''. You can
  list multiple lookups, separated by commas. For the test to end up
  with a green status, all lookups must succeed. 



 



\end{description}

\subsection{OTHER NETWORK TESTS}
\begin{description}
\item[\index{ntp}] Check for a running NTP (Network Time Protocol) server on
  this host. This test uses the ``ntpdate'' utility to check for a NTP
  server - you should either have ntpdate in your PATH, or set the
  location of the ntpdate program in \$BBHOME/etc/bbsys.local 


\item[rpc[=rpcservice1,rpcservice2,...]] Check for one or more
  available RPC services. This check is indirect in that it only
  queries the RPC Portmapper on the host, not the actual service. 


  If only ``rpc'' is given, the test only verifies that the portmapper
  is available on the remote host. If you want to check that one or
  more RPC services are registered with the portmapper, list the names
  of the desired RPC services after the equals-sign. E.g. for a
  working NFS server the ``mount'', ``nlockmgr'' and ``nfs'' services
  must be available; this can be checked with
  ``rpc=mount,nlockmgr,nfs''. 



  This test uses the rpcinfo tool for the actual test; if this tool is
  not available in the PATH of bbtest-net, you must define the RPCINFO
  environment variable to point at this tool. See
  \emph{hobbitserver.cfg(5)}


\end{description}
\subsection{HTTP TESTS}
 Simple testing of a http URL is done simply by putting the URL into
 the bb-hosts file. Note that this only applies to URL's that begin
 with ``http:'' or ``https:''. 


  The following items describe more advanced forms of http URL's. 


 \begin{description}
\item[Basic Authentication with username/password] If the URL requires
  authentication in the form of a username and password, it is most
  likely using the HTTP ``Basic'' authentication. bbtest-net support
  this, and you can provide the username and password either by
  embedding them in the URL e.g.  

 
\url{http://USERNAME:PASSWORD@www.sample.com/} 
 or by putting the username and password into the ~/.netrc file (see
 \emph{ftp(1)} for details). 


 

\item[Authentication with SSL client certificates] An SSL client
  certificate can be used for authentication. To use this, the client
  certificate must be stored in a PEM-formatted file together with the
  client certificate key, in the \$BBHOME/certs/ directory. The URL is
  then given as  

 
\url{http://CERT:FILENAME@www.sample.com/} 
 The ``CERT:'' part is literal - i.e. you write C-E-R-T-colon and then
 the filename of the PEM-formatted certificate.  

 A PEM-formatted certificate file can be generated based on
 certificates stored in Microsoft Internet Explorer and OpenSSL. Do as
 follows:  

 From the MSIE Tools-Options menu, pick the Content tab, click on
 Certificates, choose the Personal tab, select the certificate and
 click Export. Make sure you export the private key also. In the
 Export File Format, choose PKCS 12 (.PFX), check the ``Include all
 certificates'' checkbox and uncheck the ``Enable strong
 protection''. Provide a temporary password for the exported file, and
 select a filename for the PFX-file.  

 Now run ``openssl pkcs12 -in file.pfx -out file.pem''. When prompted
 for the ``Import Password'', provide the temporary password you gave
 when exporting the certificate. Then provide a ``PEM pass phrase''
 (twice) when prompted for one.  

 The file.pem file is the one you should use in the FILENAME field in
 the URL - this file must be kept in \$BBHOME/certs/. The PEM pass
 phrase must be put into a file named the same as the certificate, but
 with extension ``.pass''. E.g. if you have the PEM certificate in
 \$BBHOME/certs/client.pem, you must put the pass phrase into the
 \$BBHOME/certs/client.pass file. Make sure to protect this file with
 Unix permissions, so that only the user running Hobbit can read it. 

 

\item[Forcing an HTTP or SSL version] Some SSL sites will only allow
  you to connect, if you use specific ``dialects'' of HTTP or
  SSL. Normally this is auto-negotiated, but experience shows that
  this fails on some systems. 


  bbtest-net can be told to use specific dialects, by adding one or
  more ``dialect names'' to the URL scheme, i.e. the ``http'' or
  ``https'' in the URL: 



  * ``2'', e.g. https2://www.sample.com/ : use only SSLv2  
 * ``3'', e.g. https3://www.sample.com/ : use only SSLv3  
 * ``m'', e.g. httpsm://www.sample.com/ : use only 128-bit ciphers  
 * ``h'', e.g. httpsh://www.sample.com/ : use only $>$128-bit ciphers  
 * ``10'', e.g. http10://www.sample.com/ : use HTTP 1.0  
 * ``11'', e.g. http11://www.sample.com/ : use HTTP 1.1 


  These can be combined where it makes sense, e.g to force SSLv2 and HTTP 1.0 you would use ``https210''. 


 

\item[Testing sites by IP-address] bbtest-net ignores the ``testip''
  tag normally used to force a test to use the IP-address from the
  bb-hosts file instead of the hostname, when it performs http and
  https tests. 


  The reason for this is that it interacts badly with virtual hosts,
  especially if these are IP-based as is common with https-websites. 



  Instead the IP-address to connect to can be overridden by specifying it as: 


 \url{http://www.sample.com=1.2.3.4/index.html} 



  The ``=1.2.3.4'' will case bbtest-net to run the test against the
  IP-address ``1.2.3.4'', but still trying to access a virtual website
  with the name ``www.sample.com``. 



  The ``=ip.address.of.host'' must be the last part of the hostname, so if you need to combine this with e.g. an explicit portnumber, it should be done as 


 \url{http://www.sample.com:3128=1.2.3.4/index.html} 



 

\item[HTTP Testing via proxy] bbtest-net supports the Big Brother
  syntax for specifying an HTTP proxy to use when performing http
  tests. This syntax just joins the proxy- and the target-URL into
  one, e.g.  

 
\url{http://webproxy.sample.com:3128/http://www.foo.com/} 
 would be the syntax for testing the www.foo.com website via the proxy
 running on ``webproxy.sample.com'' port 3128. 


  If the proxy portnumber is not specified, the default HTTP portnumber (80) is used. 


  If your proxy requires authentication, you can specify the username and password inside the proxy-part of the URL, e.g.  
 
\url{http://fred:Wilma1@webproxy.sample.com:3128/http://www.foo.com/}
will authenticate to the proxy using a username of ``fred'' and a
password of ``Wilma1'', before requesting the proxy to fetch the
www.foo.com homepage. 



  Note that it is not possible to test https-sites via a proxy, nor is it possible to use https for connecting to the proxy itself. 


\item[cont[=COLUMN];URL;[expected\_data\_regexp|\#digesttype:digest]]
  This tag is used to specify a http/https check, where it is also
  checked that specific content is present in the server response. 


  If the URL itself includes a semi-colon, this must be escaped as
  '\%3B' to avoid confusion over which semicolon is part of the URL,
  and which semicolon acts as a delimiter. 



  The data that must be returned can be specified either as a regular
  expression (except that $<$space$>$ is not allowed) or as a message
  digest (typically using an MD5 sum or SHA-1 hash). 



  The regex is pre-processed for backslash ``$\backslash$'' escape
  sequences. So you can really put any character in this string by
  escaping it first:  

 
$\backslash$nNewline(LF,ASCII10decimal)  
 
$\backslash$rCarriagereturn(CR,ASCII13decimal)  
 
$\backslash$tTAB(ASCII8decimal)  
 
$\backslash$$\backslash$Backslash(ASCII92decimal)  
 
$\backslash$XXThecharacterwithASCIIhex-valueXX  



  If you must have whitespace in the regex, use the [[:space:]]
  syntax, e.g. if you want to test for the string ``All is OK'', use
  ``All[[:space:]]is[[:space:]]OK''. Note that this may depend on your
  particular implementation of the regex functions found in your C
  library. Thanks to Charles Goyard for this tip. 



  Note: If you are migrating from the ``cont2.sh'' script, you must
  change the '\_' used as wildcards by cont2.sh into '.' which is the
  regular-expression wildcard character. 



  Message digests can use different algorithms: ``md5'', ``sha1'',
  ``sha256'', ``sha512'' or ``rmd160''. The digest is calculated on
  the data portion of the response from the server, i.e. HTTP headers
  are not included in the digest (as they change from one request to
  the next). Note that the md5 and sha1 hash algorithms are no longer
  considered cryptographically safe. 



  The expected digest value can be computed with the \emph{bbdigest(1)}
 utility. 


  ``cont'' tags in bb-hosts result in two status reports: One status
  with the ``http'' check, and another with the ``content'' check. 



  As with normal URL's, the extended syntax described above can be
  used e.g. when testing SSL sites that require the use of SSLv2 or
  strong ciphers. 



  The column name for the result of the content check is by default
  called ``content'' - you can change the default with the
  ``--content=NAME'' option to bbtest-net. See \emph{bbtest-net(1)}
  for a description of this option. 



  If more than one content check is present for a host, the first
  content check is reported in the column ``content'', the second is
  reported in the column ``content1'', the third in ``content2'' etc. 



  You can also specify the columnname directly in the test
  specification, by writing it as
  ``cont=COLUMN;\url{http://...}``. Column-names cannot include
  whitespace or semi-colon. 



  The content-check status by default includes the full URL that was
  requested, and the HTML data returned by the server. You can hide
  the HTML data on a per-host (not per-test) basis by adding the
  \textbf{HIDEHTTP} tag to the host entry. 



 

\item[content=URL] This syntax is deprecated. You should use the
  ``cont'' tag instead, see above. 


 

\item[post[=COLUMN];URL;form-data;[expected\_data\_regexp|\#digesttype:digest]]
  This tag can be used to test web pages, that use an input form. Data
  can be posted to the form by specifying them in the form-data field,
  and the result can be checked as if it was a normal content check
  (see above for a description of the cont-tag and the restrictions on
  how the URL must be writen). 


  The form-data field by default is in the
  ``application/x-www-form-urlencoded'' format, which is the most
  commonly used format for web forms. 



  E.g. if you have a web form defined like this: 


  
$<$formaction=''/cgi-bin/form.cgi''method=''post''$>$  
 
$<$p$>$Givenname$<$inputtype=''text''name=''givenname''$>$$<$/p$>$  
 
$<$p$>$Surname$<$inputtype=''text''name=''surname''$>$$<$/p$>$  
 
$<$inputtype=''submit''value=''Send''$>$  
 
$<$/form$>$ 


  and you want to post the value ``John'' to the first field and ``Doe
  Jr.'' to the second field, then the formdata field would be 



  
givenname=John\&surname=Doe$\backslash$x20Jr. 


  Note that any spaces in the input value is replaced with '+'. 


  If your form-data requires a different content-type, you can specify
  it by beginning the form-data with \textbf{(content-type=TYPE)},
  e.g. ``(content-type=text/xml)'' followed by the POST data. Note
  that as with normal forms, the POST data should be specified using
  escape-sequences for reserved characters: ``space'' should be
  entered as ``$\backslash$x20'', double quote as ``$\backslash$x22'',
  newline as ``$\backslash$n'', carriage-return as ``$\backslash$r'',
  TAB as ``$\backslash$t'', backslash as
  ``$\backslash$$\backslash$''. Any byte value can be entered using
  ``$\backslash$xNN'' with NN being the hexadecimal value,
  e.g. ``$\backslash$x20'' is the space character. 



  The [expected\_data\_regexp|\#digesttype:digest] is the expected
  data returned from the server in response to the POST. See the
  ``cont;'' tag above for details. If you are only interested in
  knowing if it is possible to submit the form (but don't care about
  the data), this can be an empty string - but the ';' at the end is
  required. 



 

\item[nocont[=COLUMN];URL;forbidden\_data\_regexp] This tag works just
  like ``cont'' tag, but reverses the test. It is green when the
  ``forbidden\_data\_regexp'' is NOT found in the response, and red
  when it IS found. So it can be used to watch for data that should
  NOT be present in the response, e.g. a server error message. 


 

\item[nopost[=COLUMN];URL;form-data;expected\_data\_regexp] This tag
  works just like ``post'' tag, but reverses the test. It is green
  when the ``forbidden\_data\_regexp'' is NOT found in the response,
  and red when it IS found. So it can be used to watch for data that
  should NOT be present in the response, e.g. a server error message. 


 

\item[type[=COLUMN];URL;expected\_content\_type] This is a variant of
  the content check - instead of checking the content data, it checks
  the type of the data as given by the HTTP Content-Type: header. This
  can used to check if a URL returns e.g. a PDF file, regardless of
  what is inside the PDF file. 


 

\item[httpstatus[=COLUMN];URL;okstatusexpr;notokstatusexpr] This is
  used to explicitly test for certain HTTP statuscodes returned when
  the URL is requested. The \textbf{okstatusexpr} and
  \textbf{nokokstatusexpr} expressions are Perl-compatible regular
  expressions, e.g. ``2..|302'' will match all OK codes and the
  redirect (302) status code. If the URL cannot be retrived, the
  status is ``999''. 


 

\item[\index{HIDEHTTP}] The status display for HTTP checks usually includes
  the URL, and for content checks also the actual data from the
  webpage. If you would like to hide these from view, then the
  HIDEHTTP tag will keep this information from showing up on the
  status webpages. 


 

\item[browser=BROWSERNAME] By default, Hobbit sends an HTTP
  ``User-Agent'' header identifying it a ``Hobbit''. Some websites
  require that you use a specific browser, typically Internet
  Explorer. To cater for testing of such sites, this tag can be used
  to modify the data sent in the User-Agent header.   E.g. to perform
  an HTTP test with Hobbit masquerading as an Internet Explorer 6.0
  browser, use \textbf{browser=''Mozilla/4.0 (compatible; MSIE 6.0;
    Windows NT 5.0)''}. If you do not know what the User-Agent header
  should be, open up the browser that works with this particular site,
  and open the URL
  ``javascript:document.writeln(navigator.userAgent)'' (just copy this
  into the ``Open URL'' dialog. The text that shows up is what the
  browser sends as the User-Agent header. 


 


\end{description}

\subsection{LDAP (DIRECTORY SERVER) TESTS}
\begin{description}
\item[\index{ldap}]
\item[\index{ldaps}] Simple check for an LDAP service. This check merely looks
  for any service running on the ldap/ldaps service port, but does not
  perform any actual LDAP transaction. 

\item[ldap://hostport/dn[?attrs[?scope[?filter[?exts]]]]] Check for an
  LDAP service by performing an LDAP request. This tag is in the form
  of an LDAP URI (cf. RFC 2255). This type of LDAP test requires that
  \emph{bbtest-net(1)} was built with support for LDAP, e.g. via the
  OpenLDAP library. The components of the LDAP URI are:
  \begin{verbatim}


  \emph{hostport}
 is a host name with an optional ":portnumber"
  \emph{dn}
 is the search base
  \emph{attrs}
 is a comma separated list of attributes to request
  \emph{scope}
 is one of these three strings:
    base one sub (default=base)
  \emph{filter}
 is filter
  \emph{exts}
 are recognized set of LDAP and/or API extensions.

\end{verbatim}


 

\item[ldaps://hostport/dn[?attrs[?scope[?filter[?exts]]]]] LDAP
  service check using LDAPv3 and STARTTLS for talking to an LDAP
  server that requires TLS encryption. See \emph{bbtest-net(1)} for a
  discussion of the different ways of running LDAP servers with
  SSL/TLS, and which of these are supported by bbtest-net. 


 

\item[ldaplogin=username:password] Define a username and password to
  use when binding to the LDAP server for ldap URI tests. If not
  specified, bbtest-net will attempt an anonymous bind. 


 

\item[\index{ldapyellowfail}] Used with an LDAP URL test. If the LDAP query
  fails during the search of the directory, the ldap status is
  normally reported as ``red'' (alarm). This tag reduces a search
  failure to a ``yellow'' (warning) status. 


 


\end{description}
\subsection{PERFORMANCE MONITORING TESTS}
\begin{description}
\item[apache[=URL]] If you are running an Apache webserver, adding
  this tag makes \emph{bbtest-net(1)} collect performance statistics
  from the Apache webserver by querying the URL
  \textbf{\url{http://IP.ADDRESS.OF.HOST/server-status?auto}}. The
  response is sent as a data-report and processed by the Hobbit
  hobbitd\_rrd module into an RRD file and an ``apache'' graph. If
  your webserver requires e.g. authentication, or runs on a different
  URL for the server-status, you can provide the full URL needed to
  fetch the server-status page,
  e.g.
  \textbf{apache=\url{http://LOGIN:PASSWORD@10.0.0.1/server-status?auto}}
  for a password protected server-status page, or
  \textbf{apache=\url{http://10.0.0.1:8080/apache/server-status?auto}}
  for a server listening on port 8080 and with a different path to the
  server-status page. 



  Note that you need to enable the server-status URL in your Apache
  configuration. The following configuration is needed: 



  
$<$Location/server-status$>$  
 
SetHandlerserver-status  
 
Orderdeny,allow  
 
Denyfromall  
 
allowfrom127.0.0.1  
 
$<$/Location$>$  
 
ExtendedStatusOn 


  Change ``127.0.0.1'' to the IP-address of the server that runs your network tests. 


 


\end{description}
\subsection{DEFAULT HOST}
 If you have certain tags that you want to apply to all hosts, you can
 define a host name ``.default.'' and put the tags on that host. Note
 that per-host definitions will override the default ones. 


 \textbf{NOTE:}
 The ``.default.'' host entry will only accept the following tags -
 others are silently ignored: NOCOLUMNS, COMMENT, DESCR, CLASS,
 dialup, testip, nobb2, nodisp, noinfo, notrends, TRENDS, NOPROPRED,
 NOPROPYELLOW, NOPROPPURPLE, NOPROPACK, REPORTTIME, WARNPCT, NET,
 noclear, nosslcert, ssldays, DOWNTIME, depends, noping, noconn,
 trace, notrace, HIDEHTTP, browser, pulldata. Specifically, note that
 network tests, ``badTEST'' settings, and alternate pageset relations
 cannot be listed on the ``.default.'' host. 



 
\subsection{SENDING SUMMARIES TO REMOTE HOBBIT SERVERS}
\begin{description}
\item[summary ROW.COLUMN IP URL] If you have multiple Hobbit servers,
  the ``summary'' directive lets you form a hierarchy of servers by
  sending the overall status of this server to a remote Hobbit server,
  which then displays this in a special summary section. E.g. if your
  offices are spread over three locations, you can have a Hobbit
  server at each office. These branch-office Hobbits have a
  ``summary'' definition in their bb-hosts file that makes them report
  the overall status of their branch Hobbit to the central Hobbit
  server you maintain at the corporate headquarters. 


  Multiple ``summary'' definitions are allowed. 


  The ROW.COLUMN setting defines how this summary is presented on the
  server that receives the summary. The ROW text will be used as the
  heading for a summary line, and the COLUMN defines the name of the
  column where this summary is shown - like the hostname and testname
  used in the normal displays. The IP is the IP-address of the
  \textbf{remote} (upstream) Hobbit server, where this summary is
  sent). The URL is the URL of your \textbf{local} Hobbit server. 



  The URL need not be that of your Hobbit server's main page - it
  could be the URL of a subpage on the local Hobbit server. Hobbit
  will report the summary using the color of the page found at the URL
  you specify. E.g. on your corporate Hobbit server you want a summary
  from the Las Vegas office - but you would like to know both what the
  overall status is, and what is the status of the servers on the
  critical Sales department back-office servers in Las Vegas. So you
  configure the Las Vegas Hobbit server to send \textbf{two}
  summaries: 





  
summaryVegas.All10.0.1.1\url{http://vegas.foo.com/hobbit/} 
 
summaryVegas.Sales10.0.1.1\url{http://vegas.foo.com/hobbit/sales/}


 


  This gives you one summary line for Baltimore, with two columns: An
  ``All'' column showing the overall status, and a ``Sales'' column
  showing the status of the ``sales'' page on the Baltimore Hobbit
  server. 



  Note: Pages defined using alternate pageset definitions cannot be
  used, the URL must point to a webpage from the default set of Hobbit
  webpages. 



 


\end{description}
\subsection{OTHER TAGS}
\begin{description}
\item[pulldata[=[IP][:port]]] This option is recognized by the
  \emph{hobbitfetch(8)} utility, and causes it to poll the host for
  client data. The optional IP-address and port-number can be used if
  the client-side \emph{msgcache(8)} daemon is listening on a
  non-standard IP-address or port-number. 


 

\end{description}
\subsection{FILES}
\textbf{~hobbit/server/etc/bb-hosts}


 
\subsection{SEE ALSO}
bbgen(1), bbtest-net(1), bbdigest(1), hobbitserver.cfg(5), hobbit(7) 

 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{hobbit-alerts.cfg - Configuration for the hobbitd\_alert module}
 
\subsection{SYNOPSIS}
\textbf{~hobbit/server/etc/hobbit-alerts.cfg}


 
\subsection{DESCRIPTION}
 The hobbit-alerts.cfg file controls the sending of alerts by Hobbit when monitoring detects a failure. 

 
\subsection{FILE FORMAT}
 The configuration file consists of \textbf{rules}
, that may have one or more \textbf{recipients}
 associated. A recipient specification may include additional rules that limit the circumstances when this recipient is eligible for receiving an alert. 

  Blank lines and lines starting with a hash mark (\#) are treated as comments and ignored. Long lines can be broken up by putting a backslash at the end of the line and continuing the entry on the next line. 


 
\subsection{RULES}
 A rule consists of one of more filters using these keywords: 

 \textbf{PAGE=targetstring}
 Rule matching an alert by the name of the page in BB. This is the
 path of the page as defined in the bb-hosts file. E.g. if you have
 this setup: 

\begin{description}
\item[Example]

\begin{verbatim}

page servers All Servers
subpage web Webservers
10.0.0.1 www1.foo.com
subpage db Database servers
10.0.0.2 db1.foo.com

\end{verbatim}


\end{description}



  Then the ``All servers'' page is found with \textbf{PAGE=servers}
, the ``Webservers'' page is \textbf{PAGE=servers/web}
 and the ``Database servers'' page is \textbf{PAGE=servers/db}
. Note that you can also use regular expressions to specify the page name, e.g. \textbf{PAGE=\%.*/db}
 would find the ``Database servers'' page regardless of where this page was placed in the hierarchy. 


  The PAGE name of top-level page is an empty string. To match this, use \textbf{PAGE=\%\^{}\$}
 to match the empty string. 


 


 \textbf{EXPAGE=targetstring}
 Rule excluding an alert if the pagename matches. 


 \textbf{HOST=targetstring}
 Rule matching an alert by the hostname. 


 \textbf{EXHOST=targetstring}
 Rule excluding an alert by matching the hostname. 


 \textbf{SERVICE=targetstring}
 Rule matching an alert by the service name. 


 \textbf{EXSERVICE=targetstring}
 Rule excluding an alert by matching the service name. 


 \textbf{GROUP=groupname}
 Rule matching an alert by the group name. Groupnames are assigned to a status via the GROUP setting in the hobbit-clients.cfg file. 


 \textbf{EXGROUP=groupname}
 Rule excluding an alert by the group name. Groupnames are assigned to a status via the GROUP setting in the hobbit-clients.cfg file. 


 \textbf{COLOR=color[,color]}
 Rule matching an alert by color. Can be ``red'', ``yellow'', or ``purple''. The forms ``!red'', ``!yellow'' and ``!purple'' can also be used to NOT send an alert if the color is the specified one. 


 \textbf{TIME=timespecification}
 Rule matching an alert by the time-of-day. This is specified as the DOWNTIME timespecification in the bb-hosts file. 


 \textbf{DURATION$>$time, DURATION$<$time}
 Rule matcing an alert if the event has lasted longer/shorter than the given duration. E.g. DURATION$>$1h (lasted longer than 1 hour) or DURATION$<$30 (only sends alerts the first 30 minutes). The duration is specified as a number, optionally followed by 'm' (minutes, default), 'h' (hours) or 'd' (days). 


 \textbf{RECOVERED}
 Rule matches if the alert has recovered from an alert state. 


 \textbf{NOTICE}
 Rule matches if the message is a ``notify'' message. This type of message is sent when a host or test is disabled or enabled. 


  The ``targetstring'' is either a simple pagename, hostname or servicename, OR a '\%' followed by a Perl-compatible regular expression. E.g. ``HOST=\%www(.*)'' will match any hostname that begins with ``www''. The same for the ``groupname'' setting. 


 
\subsection{RECIPIENTS}
 The recipients are listed after the initial rule. The following keywords can be used to define recipients: 

 \textbf{MAIL address[,address]}
 Recipient who receives an e-mail alert. This takes one parameter, the e-mail address. 


 \textbf{SCRIPT /path/to/script recipientID}
 Recipient that invokes a script. This takes two parameters: The script filename, and the recipient that gets passed to the script. 


 \textbf{IGNORE}
 This is used to define a recipient that does NOT trigger any alerts, and also terminates the search for more recipients. It is useful if you have a rule that handles most alerts, but there is just that one particular server where you dont want cpu alerts on Monday morning. Note that the IGNORE recipient always has the STOP flag defined, so when the IGNORE recipient is matched, no more recipients will be considered. So the location of this recipient in your set of recipients is important. 


 \textbf{FORMAT=formatstring}
 Format of the text message with the alert. Default is ``TEXT'' (suitable for e-mail alerts). ``PLAIN'' is the same as text, but without the URL link to the status webpage. ``SMS'' is a short message with no subject for SMS alerts. ``SCRIPT'' is a brief message template for scripts. 


 \textbf{REPEAT=time}
 How often an alert gets repeated. As with DURATION, time is a number optionally followed by 'm', 'h' or 'd'. 


 \textbf{UNMATCHED}
 The alert is sent to this recipient ONLY if no other recipients received an alert for this event. 


 \textbf{STOP}
 Stop looking for more recipients after this one matches. This is implicit on IGNORE recipients. 


 \textbf{Rules}
 You can specify rules for a recipient also. This limits the alerts sent to this particular recipient. 


 
\subsection{MACROS}
 It is possible to use \textbf{macros}
 in the configuration file. To define a macro: 

 \$MYMACRO=text extending to end of line 



  After the definition of a macro, it can be used throughout the file. Wherever the text \$MYMACRO appears, it will be substituted with the text of the macro before any processing of rules and recipients. 


  It is possible to nest macros, as long as the macro is defined before it is used. 


 
\subsection{ALERT SCRIPTS}
 Alerts can go out via custom scripts, by using the SCRIPT keyword for a recipient. Such scritps have access to the following environment variables: 

 \textbf{BBALPHAMSG}
 The full text of the status log triggering the alert 


 \textbf{ACKCODE}
 The ``cookie'' that can be used to acknowledge the alert 


 \textbf{RCPT}
 The recipientID from the SCRIPT entry 


 \textbf{BBHOSTNAME}
 The name of the host that the alert is about 


 \textbf{MACHIP}
 The IP-address of the host that has a problem 


 \textbf{BBSVCNAME}
 The name of the service that the alert is about 


 \textbf{BBSVCNUM}
 The numeric code for the service. From the SVCCODES definition. 


 \textbf{BBHOSTSVC}
 HOSTNAME.SERVICE that the alert is about. 


 \textbf{BBHOSTSVCCOMMAS}
 As BBHOSTSVC, but dots in the hostname replaced with commas 


 \textbf{BBNUMERIC}
 A 22-digit number made by BBSVCNUM, MACHIP and ACKCODE. 


 \textbf{RECOVERED}
 Is ``1'' if the service has recovered. 


 \textbf{EVENTSTART}
 Timestamp when the current status (color) began. 


 \textbf{SECS}
 Number of seconds the service has been down. 


 \textbf{DOWNSECSMSG}
 When recovered, holds the text ``Event duration : N'' where N is the DOWNSECS value. 


 \textbf{CFID}
 Line-number in the hobbit-alerts.cfg file that caused the script to be invoked. Can be useful when troubleshooting alert configuration rules. 


 
\subsection{SEE ALSO}
hobbitd\_alert(8), hobbitd(8), hobbit(7), the ``Configuring Hobbit Alerts'' guide in the Online documentation. 

 


  

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{hobbit-clients.cfg - Configuration file for the hobbitd\_client module}
 
\subsection{SYNOPSIS}
\textbf{~hobbit/server/etc/hobbit-clients.cfg}


 
\subsection{DESCRIPTION}
 The hobbit-clients.cfg file controls what color is assigned to the status-messages that are generated from the Hobbit client data - typically the cpu, disk, memory, procs- and msgs-columns. Color is decided on the basis of some \textbf{settings}
 defined in this file; settings apply to specific hosts through a set of \textbf{rules}
. 

  Note: This file is only used on the Hobbit server - it is not used by the Hobbit client, so there is no need to distribute it to your client systems. 


 
\subsection{FILE FORMAT}
 Blank lines and lines starting with a hash mark (\#) are treated as comments and ignored. 

 


 
\subsection{CPU STATUS COLUMN SETTINGS}


 \textbf{LOAD warnlevel paniclevel}



  If the system load exceeds ``warnlevel'' or ``paniclevel'', the ``cpu'' status will go yellow or red, respectively. These are decimal numbers. 


  Defaults: warnlevel=5.0, paniclevel=10.0 


 \textbf{UP bootlimit toolonglimit}



  The cpu status goes yellow if the system has been up for less than ``bootlimit'' time, or longer than ``toolonglimit''. The time is in minutes, or you can add h/d/w for hours/days/weeks - eg. ``2h'' for two hours, or ``4w'' for 4 weeks. 


  Defaults: bootlimit=1h, toolonglimit=-1 (infinite). 


 


 \textbf{CLOCK max.offset}



  The cpu status goes yellow if the system clock on the client differs more than ``max.offset'' seconds from that of the Hobbit server. Note that this is not a particularly accurate test, since it is affected by network delays between the client and the server, and the load on both systems. You should therefore not rely on this being accurate to more than +/- 5 seconds, but it will let you catch a client clock that goes completely wrong. The default is NOT to check the system clock.  
\textbf{NOTE:}
 Correct operation of this test obviously requires that the system clock of the Hobbit server is correct. You should therefore make sure that the Hobbit server is synchronized to the real clock, e.g. by using NTP. 


 


  Example: Go yellow if the load average exceeds 5, and red if it exceeds 10. Also, go yellow for 10 minutes after a reboot, and after 4 weeks uptime. Finally, check that the system clock is at most 15 seconds offset from the clock of the Hobbit system. \begin{description}
\item[Example]

\begin{verbatim}

LOAD 5 10
UP 10m 4w
CLOCK 15

\end{verbatim}


\end{description}



 


 
\subsection{DISK STATUS COLUMN SETTINGS}


 \textbf{DISK filesystem warnlevel paniclevel}
 \textbf{DISK filesystem IGNORE}



  If the utilization of ``filesystem'' is reported to exceed ``warnlevel'' or ``paniclevel'', the ``disk'' status will go yellow or red, respectively. ``warnlevel'' and ``paniclevel'' are either the percentage used, or the space available as reported by the local ``df'' command on the host. For the latter type of check, the ``warnlevel'' must be followed by the letter ``U'', e.g. ``1024U''. 


  The special keyword ``IGNORE'' causes this filesystem to be ignored completely, i.e. it will not appear in the ``disk'' status column and it will not be tracked in a graph. This is useful for e.g. removable devices, backup-disks and similar hardware. 


  ``filesystem'' is the mount-point where the filesystem is mounted, e.g. ``/usr'' or ``/home''. A filesystem-name that begins with ``\%'' is interpreted as a Perl-compatible regular expression; e.g. ``\%\^{}/oracle.*/'' will match any filesystem whose mountpoint begins with ``/oracle''. 


  Defaults: warnlevel=90\%, paniclevel=95\% 


 
\subsection{MEMORY STATUS COLUMN SETTINGS}


 \textbf{MEMPHYS warnlevel paniclevel}
 
\textbf{MEMACT warnlevel paniclevel}
 
\textbf{MEMSWAP warnlevel paniclevel}



  If the memory utilization exceeds the ``warnlevel'' or ``paniclevel'', the ``memory'' status will change to yellow or red, respectively. Note: The words ``PHYS'', ``ACT'' and ``SWAP'' are also recognized. 


  Example: Go yellow if more than 20\% swap is used, and red if more
  than 40\% swap is used or the actual memory utilisation exceeds
  90\%. Dont alert on physical memory usage. 

\begin{description}
\item[\index{Example}]
\begin{verbatim}

MEMSWAP 20 40
MEMACT 90 90
MEMPHYS 101 101

\end{verbatim}


\end{description}



  Defaults: \begin{description}
\item[Example]
\begin{verbatim}

MEMPHYS warnlevel=100 paniclevel=101 (i.e. it will never go red).
MEMSWAP warnlevel=50 paniclevel=80
MEMACT  warnlevel=90 paniclevel=97

\end{verbatim}


\end{description}



 


 \textbf{PAGING warnlevel paniclevel}



  Paging rates (Page in \& out per second) for operating systems that measure that data (currently only z/VM and z/VME). Defaults are warnlevel=5, paniclevel=10. 


 
\subsection{PROCS STATUS COLUMN SETTINGS}


 \textbf{PROC processname minimumcount maximumcount color [TRACK=id] [TEXT=text]}



  The ``ps'' listing sent by the client will be scanned for how many processes containing ``processname'' are running, and this is then matched against the min/max settings defined here. If the running count is outside the thresholds, the color of the ``procs'' status changes to ``color''. 


  To check for a process that must NOT be running: Set minimum and maximum to 0. 


  ``processname'' can be a simple string, in which case this string must show up in the ``ps'' listing as a command. The scanner will find a ps-listing of e.g. ``/usr/sbin/cron'' if you only specify ``processname'' as ``cron''. ``processname'' can also be a Perl-compatiable regular expression, e.g. ``\%java.*inst[0123]'' can be used to find entries in the ps-listing for ``java -Xmx512m inst2'' and ``java -Xmx256 inst3''. In that case, ``processname'' must begin with ``\%'' followed by the regular expression. Note that Hobbit defaults to case-insensitive pattern matching; if that is not what you want, put ``(?-i)'' between the ``\%'' and the regular expression to turn this off. E.g. ``\%(?-i)HTTPD'' will match the word HTTPD only when it is upper-case.  
 If ``processname'' contains whitespace (blanks or TAB), you must enclose the full string in double quotes - including the ``\%'' if you use regular expression matching. E.g. 


  
PROC''\%hobbitd\_channel--channel=data.*hobbitd\_rrd''11yellow 


  or 


  
PROC''java-DCLASSPATH=/opt/java/lib''25 


  You can have multiple ``PROC'' entries for the same host, all of the checks are merged into the ``procs'' status and the most severe check defines the color of the status. 


  The optional \textbf{TRACK=id}
 setting causes Hobbit to track the number of processes found in an RRD file, and put this into a graph which is shown on the ``procs'' status display. The \textbf{id}
 setting is a simple text string which will be used as the legend for the graph, and also as part of the RRD filename. It is recommended that you use only letters and digits for the ID.  
 Note that the process counts which are tracked are only performed once when the client does a poll cycle - i.e. the counts represent snapshots of the system state, not an average value over the client poll cycle. Therefore there may be peaks or dips in the actual process counts which will not show up in the graphs, because they happen while the Hobbit client is not doing any polling. 


  The optional \textbf{TEXT=text}
 setting is used in the summary of the ``procs'' status. Normally, the summary will show the ``processname'' to identify the process and the related count and limits. But this may be a regular expression which is not easily recognizable, so if defined, the \textbf{text}
 setting string will be used instead. This only affects the ``procs'' status display - it has no effect on how the rule counts or recognizes processes in the ``ps'' output. 


  Example: Check that ``cron'' is running:  
PROC cron 



  Example: Check that at least 5 ``httpd'' processes are running, but not more than 20:  
PROC httpd 5 20 



  Defaults:  
mincount=1, maxcount=-1 (unlimited), color=''red''. 
 
Note that no processes are checked by default. 



 
\subsection{MSGS STATUS COLUMN SETTINGS}


 \textbf{LOG logfilename pattern [COLOR=color] [IGNORE=excludepattern]}



  The Hobbit client extracts interesting lines from one or more logfiles - see the \emph{client-local.cfg(5)}
 man-page for information about how to configure which logs a client should look at. 


  The \textbf{LOG}
 setting determine how these extracts of log entries are processed, and what warnings or alerts trigger as a result. 


  ``logfilename'' is the name of the logfile. Only logentries from this filename will be matched against this rule. Note that ``logfilename'' can be a regular expression (if prefixed with a '\%' character). 


  ``pattern'' is a string or regular expression. If the logfile data matches ``pattern'', it will trigger the ``msgs'' column to change color. If no ``color'' parameter is present, the default is to go ``red'' when the pattern is matched. To match against a regular expression, ``pattern'' must begin with a '\%' sign - e.g ``\%WARNING|NOTICE'' will match any lines containing either of these two words. Note that Hobbit defaults to case-insensitive pattern matching; if that is not what you want, put ``(?-i)'' between the ``\%'' and the regular expression to turn this off. E.g. ``\%(?-i)WARNING'' will match the word WARNING only when it is upper-case. 


  ``excludepattern'' is a string or regular expression that can be used to filter out any unwanted strings that happen to match ``pattern''. 


  Example: Trigger a red alert when the string ``ERROR'' appears in the ``/var/adm/syslog'' file:  
LOG /var/adm/syslog ERROR 



  Example: Trigger a yellow warning on all occurrences of the word ``WARNING'' or ``NOTICE'' in the ``daemon.log'' file, except those from the ``lpr'' system:  
LOG /var/log/daemon.log \%WARNING|NOTICE COLOR=yellow IGNORE=lpr 



  Defaults:  
color=''red'', no ``excludepattern''. 



  Note that no logfiles are checked by default. Any log data reported by a client will just show up on the ``msgs'' column with status OK (green). 


 


 
\subsection{FILES STATUS COLUMN SETTINGS}


 \textbf{FILE filename [color] [things to check] [TRACK]}



 \textbf{DIR directoryname [color] [size$<$MAXSIZE] [size$>$MINSIZE] [TRACK]}



  These entries control the status of the ``files'' column. They allow you to check on various data for files and directories. 


 \textbf{filename}
 and \textbf{directoryname}
 are names of files or directories, with a full path. You can use a regular expression to match the names of files and directories reported by the client, if you prefix the expression with a '\%' character. 


 \textbf{color}
 is the color that triggers when one or more of the checks fail. 


  The \textbf{TRACK}
 keyword causes the size of the file or directory to be tracked in an RRD file, and presented in a graph on the ``files'' status display. 


  For files, you can check one or more of the following: \begin{description}
\item[noexist] triggers a warning if the file exists. By default, a warning is triggered for files that have a FILE entry, but which do not exist. 
\item[type=TYPE] where TYPE is one of ``file'', ``dir'', ``char'', ``block'', ``fifo'', ``socket'' or ``symlink''. Triggers warning if the file is not of the specified type. 
\item[ownerid=OWNER] triggers a warning if the owner does not match what is listed here. OWNER is specified either with the numeric uid, or the user name. 
\item[groupid=GROUP] triggers a warning if the group does not match what is listed here. GROUP is specified either with the numeric gid, or the group name. 
\item[mode=MODE] triggers a warning if the file permissions are not as listed. MODE is written in the standard octal notation, e.g. ``644'' for the rw-r--r-- permissions. 
\item[size$<$MAX.SIZE and size$>$MIN.SIZE] triggers a warning it the file size is greater than MAX.SIZE or less than MIN.SIZE, respectively. For filesizes, you can use the letters ``K'', ``M'', ``G'' or ``T'' to indicate that the filesize is in Kilobytes, Megabytes, Gigabytes or Terabytes, respectively. If there is no such modifier, Kilobytes is assumed. E.g. to warn if a file grows larger than 1MB, use \textbf{size$<$1024M}
. 
\item[mtime$>$MIN.MTIME mtime$<$MAX.MTIME] checks how long ago the file was last modified (in seconds). E.g. to check if a file was updated within the past 10 minutes (600 seconds): \textbf{mtime$<$600}
. Or to check that a file has NOT been updated in the past 24 hours: \textbf{mtime$>$86400}
. 
\item[mtime=TIMESTAMP] checks if a file was last modified at TIMESTAMP. TIMESTAMP is a unix epoch time (seconds since midnight Jan 1 1970 UTC). 
\item[ctime$>$MIN.CTIME, ctime$<$MAX.CTIME, ctime=TIMESTAMP] acts as the mtime checks, but for the ctime timestamp (when the directory entry of the file was last changed, eg. by chown, chgrp or chmod). 
\item[md5=MD5SUM, sha1=SHA1SUM, rmd160=RMD160SUM, sha256=SHA256SUM, sha512=SHA512SUM] trigger a warning if the file checksum using the MD5, SHA1, RMD160, SHA-256 or SHA-512 message digest algorithms do not match the one configured here. Note: The ``file'' entry in the \emph{client-local.cfg(5)}
 file must specify which algorithm to use. 

 


\end{description}



  For directories, you can check one or more of the following: \begin{description}
\item[size$<$MAX.SIZE and size$>$MIN.SIZE] triggers a warning it the directory size is greater than MAX.SIZE or less than MIN.SIZE, respectively. Directory sizes are reported in whatever unit the \textbf{du}
 command on the client uses - often KB or diskblocks - so MAX.SIZE and MIN.SIZE must be given in the same unit. 

 


\end{description}



  Experience shows that it can be difficult to get these rules right. Especially when defining minimum/maximum values for file sizes, when they were last modified etc. The one thing you must remember when setting up these checks is that the rules describe criteria that must be met - only when they are met will the status be green. 


  So ``mtime$<$600'' means ``the difference between current time and the mtime of the file must be less than 600 seconds - if not, the file status will go red''. 


 


 
\subsection{PORTS STATUS COLUMN SETTINGS}


 \textbf{PORT criteria [MIN=mincount] [MAX=maxcount] [COLOR=color] [TRACK=id] [TEXT=displaytext]}



  The ``netstat'' listing sent by the client will be scanned for how many sockets match the \textbf{criteria}
 listed. The criteria you can use are: \begin{description}
\item[LOCAL=addr] ``addr'' is a (partial) local address specification in the format used on the output from netstat. 
\item[EXLOCAL=addr] Exclude certain local adresses from the rule. 
\item[REMOTE=addr] ``addr'' is a (partial) remote address specification in the format used on the output from netstat. 
\item[EXREMOTE=addr] Exclude certain remote adresses from the rule. 
\item[STATE=state] Causes only the sockets in the specified state to be included, ``state'' is usually LISTEN or ESTABLISHED but can be any socket state reported by the clients ``netstat'' command. 
\item[EXSTATE=state] Exclude certain states from the rule. 

\end{description}



  ``addr'' is typically ``10.0.0.1:80'' for the IP 10.0.0.1, port 80. Or ``*:80'' for any local address, port 80. Note that the Hobbit clients normally report only the numeric data for IP-adresses and port-numbers, so you must specify the port number (e.g. ``80'') instead of the service name (``www'').  
 ``addr'' and ``state'' can also be a Perl-compatiable regular expression, e.g. ``LOCAL=\%[.:](80|443)'' can be used to find entries in the netstat local port for both http (port 80) and https (port 443). In that case, portname or state must begin with ``\%'' followed by the reg.expression. 


  The socket count found is then matched against the min/max settings defined here. If the count is outside the thresholds, the color of the ``ports'' status changes to ``color''. To check for a socket that must NOT exist: Set minimum and maximum to 0. 


  The optional \textbf{TRACK=id}
 setting causes Hobbit to track the number of sockets found in an RRD file, and put this into a graph which is shown on the ``ports'' status display. The \textbf{id}
 setting is a simple text string which will be used as the legend for the graph, and also as part of the RRD filename. It is recommended that you use only letters and digits for the ID.  
 Note that the sockets counts which are tracked are only performed once when the client does a poll cycle - i.e. the counts represent snapshots of the system state, not an average value over the client poll cycle. Therefore there may be peaks or dips in the actual sockets counts which will not show up in the graphs, because they happen while the Hobbit client is not doing any polling. 


  The \textbf{TEXT=displaytext}
 option affects how the port appears on the ``ports'' status page. By default, the port is listed with the local/remote/state rules as identification, but this may be somewhat difficult to understand. You can then use e.g. ``TEXT=Secure Shell'' to make these ports appear with the name ``Secure Shell'' instead. 


  Defaults: mincount=1, maxcount=-1 (unlimited), color=''red''. Note: No ports are checked by default. 


  Example: Check that the SSH daemon is listening on port 22. Track the number of active SSH connections, and warn if there are more than 5.  
 
PORTLOCAL=\%[.:]22\$STATE=LISTEN''TEXT=SSHlistener''  
 
PORTLOCAL=\%[.:]22\$STATE=ESTABLISHEDMAX=5TRACK=sshTEXT=SSH 


 
\subsection{CHANGING THE DEFAULT SETTINGS}
 If you would like to use different defaults for the settings
 described above, then you can define the new defaults after a DEFAULT
 line. E.g. this would explicitly define all of the default settings: 

\begin{description}

\item[Example]

\begin{verbatim}

DEFAULT
        UP      1h
        LOAD    5.0 10.0
        DISK    * 90 95
        MEMPHYS 100 101
        MEMSWAP 50 80
        MEMACT  90 97

\end{verbatim}


\end{description}


 


 
\subsection{RULES TO SELECT HOSTS}
 All of the settings can be applied to a group of hosts, by preceding them with rules. A rule defines of one of more filters using these keywords (note that this is identical to the rule definitions used in the \emph{hobbit-alerts.cfg(5)}
 file). 

 \textbf{PAGE=targetstring}
 Rule matching an alert by the name of the page in
 BB. ``targetstring'' is the path of the page as defined in the
 bb-hosts file. E.g. if you have this setup: 

\begin{description}

\item[Example]
\begin{verbatim}

page servers All Servers
subpage web Webservers
10.0.0.1 www1.foo.com
subpage db Database servers
10.0.0.2 db1.foo.com

\end{verbatim}


\end{description}



  Then the ``All servers'' page is found with \textbf{PAGE=servers}
, the ``Webservers'' page is \textbf{PAGE=servers/web}
 and the ``Database servers'' page is \textbf{PAGE=servers/db}
. Note that you can also use regular expressions to specify the page name, e.g. \textbf{PAGE=\%.*/db}
 would find the ``Database servers'' page regardless of where this page was placed in the hierarchy. 


  The top-level page has a the fixed name \textbf{/}
, e.g. \textbf{PAGE=/} would match all hosts on the Hobbit
  frontpage. If you need it in a regular expression, use
  \textbf{PAGE=\%\^{}/} to avoid matching the forward-slash present in
  subpage-names. 



 \textbf{EXPAGE=targetstring}
 Rule excluding a host if the pagename matches. 


 \textbf{HOST=targetstring}
 Rule matching a host by the hostname. ``targetstring'' is either a
 comma-separated list of hostnames (from the bb-hosts file), ``*'' to
 indicate ``all hosts'', or a Perl-compatible regular
 expression. E.g. ``HOST=dns.foo.com,www.foo.com`` identifies two
 specific hosts; ``HOST=\%www.*.foo.com EXHOST=www-test.foo.com''
 matches all hosts with a name beginning with ``www'', except the
 ``www-test'' host. 



 \textbf{EXHOST=targetstring}
 Rule excluding a host by matching the hostname. 


 \textbf{CLASS=classname}
 Rule match by the client class-name. You specify the class-name for a
 host when starting the client through the ``--class=NAME'' option to
 the runclient.sh script. If no class is specified, the host by
 default goes into a class named by the operating system. 



 \textbf{EXCLASS=classname}
 Exclude all hosts belonging to ``classname'' from this rule. 


 \textbf{TIME=timespecification}
 Rule matching by the time-of-day. This is specified as the DOWNTIME time specification in the bb-hosts file. E.g. ``TIME=W:0800:2200'' applied to a rule will make this rule active only on week-days between 8AM and 10PM. 


 
\subsection{DIRECTING ALERTS TO GROUPS}
 For some tests - e.g. ``procs'' or ``msgs'' - the right group of
 people to alert in case of a failure may be different, depending on
 which of the client rules actually detected a problem. E.g. if you
 have PROCS rules for a host checking both ``httpd'' and ``sshd''
 processes, then the Web admins should handle httpd-failures, whereas
 ``sshd'' failures are handled by the Unix admins. 


  To handle this, all rules can have a ``GROUP=groupname'' setting. When a rule with this setting triggers a yellow or red status, the groupname is passed on to the Hobbit alerts module, so you can use it in the alert rule definitions in \emph{hobbit-alerts.cfg(5)}
 to direct alerts to the correct group of people. 



 
\subsection{RULES: APPLYING SETTINGS TO SELECTED HOSTS}
 Rules must be placed after the settings, e.g. \begin{description}
\item[Example]\begin{verbatim}

LOAD 8.0 12.0  HOST=db.foo.com TIME=*:0800:1600

\end{verbatim}


\end{description}



  If you have multiple settings that you want to apply the same rules
  to, you can write the rules *only* on one line, followed by the
  settings. E.g. 

\begin{description}
\item[Example]\begin{verbatim}

HOST=%db.*.foo.com TIME=W:0800:1600
        LOAD 8.0 12.0
        DISK /db  98 100
        PROC mysqld 1

\end{verbatim}


\end{description}



  will apply the three settings to all of the ``db'' hosts on
  week-days between 8AM and 4PM. This can be combined with
  per-settings rule, in which case the per-settings rule overrides the
  general rule; e.g. 

\begin{description}\item[Example]\begin{verbatim}


HOST=%.*.foo.com
        LOAD 7.0 12.0 HOST=bax.foo.com
        LOAD 3.0 8.0

\end{verbatim}


\end{description}



  will result in the load-limits being 7.0/12.0 for the
  ``bax.foo.com'' host, and 3.0/8.0 for all other foo.com hosts. 

  The entire file is evaluated from the top to bottom, and the first
  match found is used. So you should put the specific settings first,
  and the generic ones last. 
 
\subsection{NOTES}
 For the LOG, FILE and DIR checks, it is necessary also to configure
 the actual file- and directory-names in the
 \emph{client-local.cfg(5)} file. If the filenames are not listed
 there, the clients will not collect any data about these
 files/directories, and the settings in the hobbit-clients.cfg file
 will be silently ignored. 


  The ability to compute file checksums with MD5, SHA1 or RMD160
  should not be used for general-purpose file integrity checking,
  since the overhead of calculating these on a large number of files
  can be significant. If you need this, look at tools designed for
  this purpose - e.g. Tripwire or AIDE. 



  At the time of writing (april 2006), the SHA-1 and RMD160 algorithms
  are considered cryptographically safe. The MD5 algorithm has been
  shown to have some weaknesses, and is not considered strong enough
  when a high level of security is required. 



\subsection{SEE ALSO}
hobbitd\_client(8), client-local.cfg(5), hobbitd(8), hobbit(7) 

 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\include{hobbitcgi.cfg.5}
\newpage
\section{HOBBITCGI.CFG}
 hobbitcgi.cfg - Command-line parameters for the Hobbit CGI tools 

 
\subsection{SYNOPSIS}
\textbf{\$BBHOME/etc/hobbitcgi.cfg}


 
\subsection{DESCRIPTION}
\emph{hobbitcgi.cfg(1)}
 controls the commandline options passed to all of the Hobbit CGI tools through their respective shell-script wrappers. Typically the options listed here are used for system-wide configuration of the CGI utilities, e.g. to define where they read configuration files. 

  The exact set of commandline options available are described in the man-page for each of the CGI utilities. 


  The file is ``sourced'' into the shell script wrapper, so assignments to the CGI-specific variables must follow standard shell-script syntax. 


 
\subsection{SETTINGS}
\begin{description}
\item[\index{CGI\_ACKINFO\_OPTS}] Options for the \emph{hobbit-ackinfo.cgi(1)}
 utility. 

 

\item[\index{CGI\_ACK\_OPTS}] Options for the \emph{bb-ack.cgi(1)}
 utility. 

 

\item[\index{CGI\_CSVINFO\_OPTS}] Options for the \emph{bb-csvinfo.cgi(1)}
 utility. 

 

\item[\index{CGI\_DATEPAGE\_OPTS}] Options for the \emph{bb-datepage.cgi(1)}
 utility. 

 

\item[\index{CGI\_ENADIS\_OPTS}] Options for the \emph{hobbit-enadis.cgi(8)}
 utility. 

 

\item[\index{CGI\_EVENTLOG\_OPTS}] Options for the \emph{bb-eventlog.cgi(1)}
 utility. 

 

\item[\index{CGI\_FINDHOST\_OPTS}] Options for the \emph{bb-findhost.cgi(1)}
 utility. 

 

\item[\index{CGI\_HIST\_OPTS}] Options for the \emph{bb-hist.cgi(1)}
 utility. 



\item[\index{CGI\_HOBBITCOLUMN\_OPTS}] Hobbit-specific options for column documentation. This uses the \emph{bb-csvinfo.cgi(1)}
 utility with the \textbf{server/etc/columndoc.cfg}
 configuration file. 

 

\item[\index{CGI\_HOBBITCONFREPORT\_OPTS}] Options for the \emph{hobbit-confreport.cgi(1)}
 utility. 

 

\item[\index{CGI\_HOBBITGRAPH\_OPTS}] Options for the \emph{hobbitgraph.cgi(1)}
 utility. 

 

\item[\index{CGI\_HOSTGRAPHS\_OPTS}] Options for the \emph{hobbit-hostgraphs.cgi(1)}
 utility. 

 

\item[\index{CGI\_NKEDIT\_OPTS}] Options for the \emph{hobbit-nkedit.cgi(1)}
 utility. 

 

\item[\index{CGI\_NKVIEW\_OPTS}] Options for the \emph{hobbit-nkview.cgi(1)}
 utility. 

 

\item[\index{CGI\_REPLOG\_OPTS}] Options for the \emph{bb-replog.cgi(1)}
 utility. 

 

\item[\index{CGI\_REP\_OPTS}] Options for the \emph{bb-rep.cgi(1)}
 utility. 

 

\item[\index{CGI\_SNAPSHOT\_OPTS}] Options for the \emph{bb-snapshot.cgi(1)}
 utility. 

 

\item[\index{CGI\_SVCHIST\_OPTS}] Options for the \emph{hobbitsvc.cgi(1)}
 utility when used to view historical logs. Note that the ``--historical'' option must be included in this setting. 

 

\item[\index{CGI\_SVC\_OPTS}] Options for the \emph{hobbitsvc.cgi(1)}
 utility. 

 


\end{description}
\subsection{SEE ALSO}
hobbit(7), the individual CGI utility man-pages. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\include{hobbit-nkview.cfg.5}
\newpage
\section{HOBBIT-NKVIEW.CGI}

hobbit-nkview.cgi - Hobbit Critical Systems view CGI

 \subsection{SYNOPSIS}
\textbf{hobbit-nkview.cgi}


 
\subsection{DESCRIPTION}
\textbf{hobbit-nkview.cgi}
 is invoked as a CGI script via the hobbit-nkview.sh CGI wrapper. 

  hobbit-nkview.cgi matches the current critical statuses against the \emph{hobbit-nkview.cfg(5)}
 file, and generates the ``Critical Systems'' view. 


 
\subsection{RELATION TO BBNK PAGE}
 This view is a replacement for the statically generated ``bbnk'' page
 provided in versions of Hobbit prior to version 4.2. Although the
 ``bbnk'' pages are supported throughout Hobbit 4.x, it is recommended
 that You switch to the newer Critical Systems view provided by this
 CGI. 


 
\subsection{OPTIONS}
\begin{description}
\item[\index{--nkacklevel=NUMBER}] Sets the acknowledgment level for
  acknowledgments sent via the \emph{hobbit-ackinfo.cgi(1)} page. Note
  that this may be overridden by the configuration of the
  hobbit-ackinfo.cgi utility. 


 

\item[\index{--env=FILENAME}] Loads the environment defined in FILENAME before
  executing the CGI script. 


 

\item[\index{--area=NAME}] Load environment variables for a specific area. NB:
  if used, this option must appear before any --env=FILENAME option. 


 

\item[\index{--debug}] Enables debugging output. 

 


\end{description}
\subsection{ENVIRONMENT VARIABLES}
\begin{description}
\item[\index{BBHOME}] Used to locate the template files for the generated web pages. 

 

\item[\index{QUERY\_STRING}] Contains the parameters for the CGI script. 

 


\end{description}
\subsection{SEE ALSO}
hobbit-ackinfo.cgi(1), hobbit(7) 

  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{CLIENT-LOCAL.CFG}
\subsection{NAME}
 client-local.cfg - Local configuration settings for Hobbit clients 

 
\subsection{SYNOPSIS}
\textbf{~hobbit/server/etc/client-local.cfg}


 
\subsection{DESCRIPTION}
 The client-local.cfg file contains settings that are used by each
 Hobbit client when it runs on a monitored host. It provides a
 convenient way of configuring clients from a central location without
 having to setup special configuration maintenance tools on all
 clients. 


  The client-local.cfg file is currently used to configure what logfiles the client should fetch data from, to be used as the basis for the ``msgs'' status column; and to configure which files and directories are being monitored in the ``files'' status column. 


  Note that there is a dependency between the client-local.cfg file
  and the \emph{hobbit-clients.cfg(5)} file. When monitoring e.g. a
  logfile, you must first enter it into the client-local.cfg file, to
  trigger the Hobbit client into reporting any data about the
  logfile. Next, you must configure hobbit-clients.cfg so the Hobbit
  server knows what to look for in the file data sent by the
  client. So: client-local.cfg defines what raw data is collected by
  the client, and hobbit-clients.cfg defines how to analyze them. 



 
\subsection{PROPAGATION TO CLIENTS}
 The client-local.cfg file resides on the Hobbit server. 

  When clients connect to the Hobbit server to send in their client
  data, they will receive part of this file back from the Hobbit
  server. The configuration received by the client is then used the
  next time the client runs. 



  This method of propagating the configuration means that there is a
  delay of up to two poll cycles (i.e. 5-10 minutes) from a
  configuration change is entered into the client-local.cfg file, and
  until you see the result in the status messages reported by the
  client. 



 
\subsection{FILE FORMAT}
 The file is divided into sections, delimited by ``[name]'' lines. A
 section name can be either an operating system identifier - linux,
 solaris, hp-ux, aix, freebsd, openbsd, netbsd, darwin - or a
 hostname. When deciding which section to send to a client, Hobbit
 will first look for a section named after the hostname of the client;
 if such a section does not exist, it will look for a section named by
 the operating system of the client. So you can configure special
 configurations for individual hosts, and have a default configuration
 for all other hosts of a certain type. 


  Apart from the section delimiter, the file format is free-form, or
  rather it is defined by the tools that make use of the
  configuration. 



 
\subsection{LOGFILE CONFIGURATION ENTRIES}
 A logfile configuration entry looks like this: 

  
log:/var/log/messages:10240  
 
ignoreMARK  
 
triggerOops 


  The \textbf{log:FILENAME:SIZE} line defines the filename of the log,
  and the maximum amount of data (in bytes) to send to the Hobbit
  server. FILENAME is usually an explicit full-path filename on the
  client. If it is enclosed in backticks, it is a command which the
  Hobbit client runs and each line of output from this command is then
  used as a filename. This allows scripting which files to monitor,
  e.g. if you have logfiles that are named with some sort of
  timestamp. 



  The \textbf{ignore PATTERN} line (optional) defines lines in the
  logfile which are ignored entirely, i.e. they are stripped from the
  logfile data before sending it to the Hobbit server. It is used to
  remove completely unwanted ``noise'' entries from the logdata
  processed by Hobbit. ``PATTERN'' is a regular expression. 



  The \textbf{trigger PATTERN}
 line (optional) is used only when there is more data in the log than
 the maximum size set in the ``log:FILENAME:SIZE'' line. The
 ``trigger'' pattern is then used to find particularly interesting
 lines in the logfile - these will always be sent to the Hobbit
 server. After picking out the ``trigger'' lines, any remaining space
 up to the maximum size is filled in with the most recent entries from
 the logfile. ``PATTERN'' is a regular expression. 



 
\subsection{COUNTING LOGENTRIES}
 A special type of log-handling is possible, where the number of lines
 matching a regular expressions are merely counted. This is
 \textbf{linecount:FILENAME}, followed by a number of lines of the
 form \textbf{ID:PATTERN}

. E.g. 

  
linecount:/var/log/messages  
 
diskerrors:I/Oerror.*device.*hd  
 
badlogins:Failedlogin 


 


 
\subsection{FILE CONFIGURATION ENTRIES}
 A file monitoring entry is used to watch the meta-data of a file:
 Owner, group, size, permissions, checksum etc. It looks like this: 


  
\url{file:/var/log/messages}[:HASH] 


  The \textbf{\url{file:FILENAME}}
 line defines the filename of the file to monitor. As with the
 ``log:'' entries, a filename enclosed in backticks means a command
 which will generate the filenames dynamically. The optional [:HASH]
 setting defines what type of hash to compute for the file:
 \textbf{md5}

, \textbf{sha1}
, \textbf{rmd160}
, \textbf{sha256}
 or \textbf{sha512}
. By default, no hash is calculated.  
\textbf{NOTE:}
 If you want to check multiple files using a wildcard, you \textbf{must}
 use a command to generate the filenames. Putting wildcards directly into the \textbf{file:}
 entry will not work. 


 
\subsection{DIRECTORY CONFIGURATION ENTRIES}
 A directory monitoring entry is used to watch the size of a directory and any sub-directories. It looks like this: 

  
dir:DIRECTORYNAME 


  The \textbf{dir:DIRECTORYNAME}
 line defines the filename of the file to monitor. As with the ``log:'' entries, a filename enclosed in backticks means a command which will generate the filenames dynamically. The Hobbit client will run the \emph{du(1)}
 command with the directoryname as parameter, and send the output back to the Hobbit server.  
\textbf{NOTE:}
 If you want to check multiple directories using a wildcard, you \textbf{must}
 use a command to generate the directory names. Putting wildcards directly into the \textbf{dir:}
 entry will not work. E.g. use something like  
dir:`find /var/log -maxdepth 1 -type d` 



  The ``du'' command used can be configured through the \textbf{DU}
 environment variable. On some systems, by default \textbf{du}
 reports data in disk blocks instead of KB (e.g. Solaris). So you may want to configure the Hobbit client to use a \textbf{du}
 command which reports data in KB, e.g. by setting  
 
DU=''du-k''  
 in the hobbitclient.cfg file. 


 
\subsection{NOTES}
 The ability of the Hobbit client to calculate file hashes and monitor
 those can be used for file integrity validation on a small
 scale. However, there is a significant processing overhead in
 calculating these every time the Hobbit client runs, so this should
 not be considered a replacement for host-based intrusion detection
 systems such as Tripwire or AIDE. 


  Use of the directory monitoring on directory structures with a large number of files and/or sub-directories can be quite ressource-intensive. 


 
\subsection{SEE ALSO}
hobbit-clients.cfg(5), hobbitd\_client(8), hobbitd(8), hobbit(7) 
