\chapter{Preface}
\label{chap:preface}



\section{Skill set  Requirements }
There are a few requirements in order for you to particpate in
hobbitmon developement.

But if you don't have these skills currently, that is fine. If you
study hard with this book. You will gain all these skills once you
finish reading this book.


\begin{itemize}
\item You are doing this for free.
\item You are doing this for free without direct pay from the
  community. And to this date not many people can make money directly.

\item Ok, you have been warned. The first requirement is C programming
  language. but it doesn't stop here only at C language. There are
  many other language like Perl, Python and shell language used to
  develop modules. 
\item subversion source code management system.
\item GNU Make.
\item GNU Autotool.
\item GNU TWW CPAM tool.
\item Latex.
\item Doxygen.
\item Editor. Emacs preferred ;)
 
\end{itemize}

\section{This book is a work in progress}

I am writing this  book about hobbit from perspective of manpages.  

I am releasing this Hobbit RTFM book while I am still writing it, in the hope that
it will prove useful to others.  I also hope that readers will contribute as they see fit.

\section{Hobbitmon Documentation Sets}


\begin{itemize}
\item
  \href{http://hobbitmon.svn.sourceforge.net/viewvc/hobbitmon/branches/tjyang/src/books/hobbitmon-developer/en/}{Hobbitmon
    Developer Guide}, this book.
\item
  \href{http://hobbitmon.svn.sourceforge.net/viewvc/hobbitmon/branches/tjyang/src/books/hobbitmon-faq-howto/en/}{Hobbitmon
    FAQs and Howto}.
\item
  \href{http://hobbitmon.svn.sourceforge.net/viewvc/hobbitmon/branches/tjyang/src/books/hobbitmon-user-guide/en/}{Hobbitmon
    User Guide}.
\item
  \href{http://hobbitmon.svn.sourceforge.net/viewvc/hobbitmon/branches/tjyang/src/books/hobbitmon-sourcecode-guide/en/}{Hobbitmon
Source code Guide}.
\item
  \href{http://hobbitmon.svn.sourceforge.net/viewvc/hobbitmon/branches/tjyang/src/hobbitmon-slides/en/}{Hobbitmon
    Slides}.

\end{itemize}

\section{Revision History}

\begin{itemize}
\item T.J. Yang:
 \begin{enumerate}
  \item Using \href{http://hg.serpentine.com/mercurial/book/}{Mercuial Book } as a template to have a quick start on
    writing this book.
 \end{enumerate}

\end{itemize}

\section{Colophon---this book is Free}

This book is licensed under the Open Publication License, and is
produced entirely using Free Software tools.  It is typeset with
\LaTeX{}; illustrations are drawn and rendered with
\href{http://www.inkscape.org/}{Inkscape}.

The complete source code for this book is published as a SVN
repository, at
\href{http://hobbitmon.svn.sourceforge.net/viewvc/hobbitmon/branches/tjyang/src/books/hobbitmon-developer/en/}{Hobbitmon
  Developer Book on sourceforge.net}.


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "00book"
%%% End: 
