\title{Upgrading from Big Brother to Hobbit}
\section{Upgrading from Big Brother to Hobbit}


 First, you should realize that this is not a fully automated
 proces. You will need to do some work yourself - especially with the
 handling of alerts.

\subsubsection{First step: Install Hobbit}


 To begin, install Hobbit as described in the Hobbit installation
 guide. I recommend that you configure Hobbit to use the same user-ID
 as your current Big Brother installation, but have it use a different
 directory for the server- and data-files. The default is to use
 \emph{~/server} and \emph{~/data} respectively, which is unlikely to
 clash with the directories where you have Big Brother installed. If
 you do need to change the directories, you must edit the top-level
 Makefile and change the BBHOME and BBVAR settings near the top of the
 file.

\subsubsection{Step two: Move the configuration files}


 A couple of configuration files can be copied directly from Big
 Brother to Hobbit:

\begin{itemize}

\item The \textbf{bb-hosts}
 file 
\item The \textbf{bb-services} file. You need only copy this if you
have used bbgen before, and added custom network tests to the
bb-services file.
\item The \textbf{cookies} file. You may not have this file - it is
only present if you have used bbgen before and have setup HTTP tests
that require cookies.

\item The \textbf{bbcombotests.cfg} file. You may not have this file -
it is only present if you have used bbgen before and have setup the
bbcombotest tool.


\end{itemize}


 The \emph{bbwarnrules.cfg}
 and \emph{bbwarnsetup.cfg} files cannot be copied over. Hobbit uses a
 very different configuration file for the alert configuration, so you
 will have to re-write your alert configuration for Hobbit. See the
 Hobbit alert configuration to learn how Hobbit alerts are configured.


 Any server-side extension-scripts can be copied from the
 \$BBHOME/ext/ directory to the ~/server/ext/ directory. You must also
 add entries for them to the Hobbit \emph{hobbitlaunch.cfg}
 file. Beware that many scripts rely on environment variables that Big
 Brother defines, but which Hobbit does not define - in that case, you
 need to setup those environment variables in the hobbitserver.cfg
 file. It is probably easiest to save this until you start running
 Hobbit, and can look at any error-output from the scripts.



 If you have modified the webpage header- and footer-files in
 \$BBHOME/web/ then you can copy the modified files over directly to
 the ~/server/web/ directory. Note that Hobbit has a number of header-
 and footer-files for the various CGI scripts that are not present in
 Big Brother, so you may need to setup a few extra files to get a
 consistent look across your new Hobbit installation.


\subsubsection{Step three: Stop Big Brother}


 You are now going to move over the data files. To avoid confusion
 about files being updated by Big Brother while they are being moved
 over to Hobbit, I recommend that you stop Big Brother now.

\subsubsection{Step four: Move the history logs}


 You may want to save the historical logfiles and the history of your
 status changes. To do that, move all of the files or directories in
 the \emph{\$BBVAR/hist/} to the \emph{~/data/hist/} directory, and
 all of the files or directories in \emph{\$BBVAR/histlogs/} to the
 \emph{~/data/histlogs/}  directory. If you prefer to keep them in the
 Big Brother directory, you can copy them over with ``cp -r'' or
 ``tar'' instead of moving them.

\subsubsection{Step five: Move the RRD files}


 The RRD files are used to generate the graphs, if you have installed
 the LARRD add-on to Big Brother. Hobbit has RRD support built-in, and
 it is obviously nice to keep the historical data that has been
 collected over time.



 The filesystem layout of the RRD files is different from Big
 Brother+LARRD to Hobbit. Instead of having all of the RRD files in
 one big directory, there is a subdirectory for each host holding only
 the RRD files for data from that host. This is easier to manage, and
 also speeds up the graph generation when you have many
 hosts. Unfortunately, it makes migrating from Big Brother to Hobbit
 slightly more complicated.



 In the Hobbit source-tree, you will find a script
 \emph{hobbitd/moverrd.sh}. This script moves or copies the RRD files
 from the Big Brother+LARRD structure into the Hobbit structure. You
 must edit a couple of settings at the beginning of the file,
 especially to set the correct directory where Big Brother stores your
 current RRD files (the SRCDIR setting). By default the script
 \emph{copies} the files over to the new structure, if you would
 rather just move them then change to ``OP'' setting to ``mv''.


 After setting up the script, run it and it should copy all of the
 RRD-files that relate to a host currently in the bb-hosts file to the
 new directory structure.

\subsubsection{Step 6: Start Hobbit}


 Start Hobbit with the ~/server/hobbit.sh start command. Look at the
 logfiles in the /var/log/hobbit directory (or elsewhere, if you did
 not choose the default logfile directory when configuring Hobbit) and
 fix any problems that show up.



 Look at the webpages generated. For the first few minutes, there will
 be some missing columns and icons for each host, since it takes some
 time for all of the tests to report a status to the new Hobbit
 daemon. After 5-10 minutes all of the standard tests should appear.


